---
title: "McMurdo Dry Valleys Nutrient Limitation"
subtitle: "  \n Processing of Biogeochemical Data"
author: "Adam J. Solon"
date: "`r Sys.Date()`"
#output: html_document
output: 
  pdf_document:
    toc: TRUE
    fig_width: 7
    fig_height: 6
    fig_caption: true
    keep_md: true
fontsize: 12pt
#editor_options: 
#  chunk_output_type: console
---
  
# Script Summary  
This script processes geochemical, biological, geographic, and physical data measured from sediments collected in Taylor Valley of the McMurdo Dry Valleys, Antarctica, and subsets samples from the file that pertain to samples which contain both soil chemistry and biology for downstream analyses. 

### Steps of this pipeline:  
1.  Create and organize directories
2.  Load R packages
3.  Input files
4.  Format Files
  a. molar conversions of percent values
  b. numeric and factor classifications
  c. convert dna concentration (Qubit) into ng per g soil
  d. re-label samples IDs with more coherent names
5.  Exploratory Data Analysis
  a. 
6.  Save Files

```{r echo = FALSE, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(include = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

```

```{r, echo = FALSE, include = FALSE}
# Change identifiers to your system and file naming. 
user <- "F:"
directory <- "/Projects"
project <- "/CryoHoles/Studies"
study <- "/Soils-Nutrient-Limitation_Solon-et-al"

sub.directory.1 <- "/data"
sub.directory.1.1 <- "/processed_files"
sub.directory.1.1.1 <- "/soil_chemistry"
sub.directory.1.1.2 <- "/dna_seq"

sub.directory.2 <- "/analyses"
sub.directory.2.1 <- "/biogeochemistry"
sub.directory.2.1.1 <- "/input"
sub.directory.2.1.2 <- "/exploratory_data_analysis"
sub.directory.2.1.2.1 <- "/histograms"
sub.directory.2.1.2.2 <- "/density"
sub.directory.2.1.2.3 <- "/outliers"
sub.directory.2.1.2.4 <- "/scatter"

```

### Set pathways and create directories  

```{r set paths for pipeline}
# First define the project and project directories. 

# Create pathway for pipeline
###################################################
path.fp <- paste0(user, directory, project, study)
if (!dir.exists(path.fp)) dir.create(path.fp)

```

```{r set paths for processed files}
#sub.directory.1 data
###################################################
sub.directory.1.fp <- paste0(path.fp, sub.directory.1)
if (!dir.exists(sub.directory.1.fp)) dir.create(sub.directory.1.fp)

# Create sub-directory  processed files
###################################################
sub.directory.1.1.fp <- paste0(sub.directory.1.fp, sub.directory.1.1)
if (!dir.exists(sub.directory.1.1.fp)) dir.create(sub.directory.1.1.fp)

# Create sub-directory soil chemistry
###################################################
sub.directory.1.1.1.fp <- paste0(sub.directory.1.1.fp, sub.directory.1.1.1)
if (!dir.exists(sub.directory.1.1.1.fp)) dir.create(sub.directory.1.1.1.fp)

# Create sub-directory dna seq
###################################################
sub.directory.1.1.2.fp <- paste0(sub.directory.1.1.fp, sub.directory.1.1.2)
if (!dir.exists(sub.directory.1.1.2.fp)) dir.create(sub.directory.1.1.2.fp)

```

```{r set paths for analyses}
#sub.directory.2 analyses
###################################################
sub.directory.2.fp <- paste0(path.fp, sub.directory.2)
if (!dir.exists(sub.directory.2.fp)) dir.create(sub.directory.2.fp)

# Create sub-directory  biogeochemistry
###################################################
sub.directory.2.1.fp <- paste0(sub.directory.2.fp, sub.directory.2.1)
if (!dir.exists(sub.directory.1.1.fp)) dir.create(sub.directory.1.1.fp)

# Create sub-directory exploratory data analysis
###################################################
sub.directory.2.1.1.fp <- paste0(sub.directory.2.1.fp, sub.directory.2.1.1)
if (!dir.exists(sub.directory.2.1.1.fp)) dir.create(sub.directory.2.1.1.fp)

# Create sub-directory exploratory data analysis
###################################################
sub.directory.2.1.2.fp <- paste0(sub.directory.2.1.fp, sub.directory.2.1.2)
if (!dir.exists(sub.directory.2.1.2.fp)) dir.create(sub.directory.2.1.2.fp)

# Create sub-directory histograms
###################################################
sub.directory.2.1.2.1.fp <- paste0(sub.directory.2.1.2.fp, sub.directory.2.1.2.1)
if (!dir.exists(sub.directory.2.1.2.1.fp)) dir.create(sub.directory.2.1.2.1.fp)

# Create sub-directory density
###################################################
sub.directory.2.1.2.2.fp <- paste0(sub.directory.2.1.2.fp, sub.directory.2.1.2.2)
if (!dir.exists(sub.directory.2.1.2.2.fp)) dir.create(sub.directory.2.1.2.2.fp)

# Create sub-directory outliers
###################################################
sub.directory.2.1.2.3.fp <- paste0(sub.directory.2.1.2.fp, sub.directory.2.1.2.3)
if (!dir.exists(sub.directory.2.1.2.3.fp)) dir.create(sub.directory.2.1.2.3.fp)

# Create sub-directory scatter
###################################################
sub.directory.2.1.2.4.fp <- paste0(sub.directory.2.1.2.fp, sub.directory.2.1.2.4)
if (!dir.exists(sub.directory.2.1.2.4.fp)) dir.create(sub.directory.2.1.2.4.fp)

```

### Load R packages  
```{r Install and load packages, echo = FALSE, include = FALSE}
# install.packages("tidyverse")
# install.packages("knitr")
library(knitr); packageVersion("knitr")
library(tidyverse); packageVersion("tidyverse")
library(moments); packageVersion("moments")

```

* r version: `r getRversion()`
* RStudio version: `r rstudioapi::versionInfo()$version`
* r packages:  
  
  knitr, `r packageVersion("knitr")`  
  tidyverse, `r packageVersion("tidyverse")`  
  moments, `r packageVersion("moments")`  

### Input Files
Required input files:   
  
1.  The 'mapping file' w/ physical, geographic, chemical, and biological data for each sample 
 
Input files and re-format for use in pipeline

```{r input data files}
# input data files
# TOC and TN data
File.1.fp <- paste0(sub.directory.1.1.1.fp, "/MDV-soil-nutrient-limitation-TOC-TN-study-comparisons.csv") 
# TP and TOP
File.2.fp <- paste0(sub.directory.1.1.1.fp, "/MDV-soil-nutrient-limitation-TP-TOP-study-comparisons.csv") 

#map file
mappingFile.fp <- paste0(sub.directory.1.1.2.fp, "/MDV-soil-nutrient-limitation-map-file.txt") 

#input data
df.cn <- read.csv(File.1.fp, header = T)
df.p <- read.csv(File.2.fp, header = T)
m <- read.table(mappingFile.fp, header = T, sep = "\t")
  
```

### Format Files  
```{r format flie toc-tn}
#format dataframe
df.cn.1 <- df.cn %>%
  mutate(across(c(SampleID, Study, Location, sample_type, Substrate), as.factor)) %>%
  mutate(umol.C_g.soil = (C_perc / 100) * (1e6 / 12.01)) %>%
  mutate(umol.N_g.soil = (N_perc / 100) * (1e6 / 14.01)) %>%
  mutate(C_to_N = as.numeric(umol.C_g.soil / umol.N_g.soil)) %>%
  mutate(across(c(umol.N_g.soil, d15N), ~ signif(.x, 2))) %>%  # Keep these at 2 significant digits
  mutate(across(c(umol.C_g.soil, C_to_N, d13C), ~ round(.x, 1)))  # Round these to 1 decimal place

#rownames as sample IDs
rownames(df.cn.1) <- df.cn.1$SampleID

```

```{r format file tp-top}
#format dataframe
df.p.1 <- df.p %>%
  mutate(SampleID = case_when(
    str_detect(samples, "MDV_FX_bulk_") ~ str_replace(samples, "MDV_FX_bulk_", "FXB_0"),
    str_detect(samples, "MDV_HR_bulk_") ~ str_replace(samples, "MDV_HR_bulk_", "HRB_0"),
    str_detect(samples, "MDV_CAN_bulk_") ~ str_replace(samples, "MDV_CAN_bulk_", "CAB_0"),
    str_detect(samples, "MDV_COM_bulk_") ~ str_replace(samples, "MDV_COM_bulk_", "COB_0"),
    str_detect(samples, "MDV_TAY_bulk_") ~ str_replace(samples, "MDV_TAY_bulk_", "TAB_0"),
    TRUE ~ samples  # Preserve original values if no match is found
  )) %>%
  mutate(across(c(SampleID, Study, Location, Substrate), as.factor)) %>%  
  mutate(umol.TP_g.soil = (p_unburned_perc / 100) * (1e6 / 30.97)) %>%  # Convert % to µmol/g using atomic mass of P
  mutate(umol.TIP_g.soil = (p_burned_perc / 100) * (1e6 / 30.97)) %>%
  mutate(umol.TOP_g.soil = (p_organic_perc / 100) * (1e6 / 30.97)) %>%
  mutate(across(c(umol.TP_g.soil, umol.TIP_g.soil), ~ signif(.x, 3))) %>%  # Round TP and TIP to 3 sig digits
  mutate(umol.TOP_g.soil = signif(umol.TOP_g.soil, 4)) %>%  # Round TOP to 4 sig digits
  dplyr::select(SampleID, everything())  # Moves SampleID to the first column

#rownames as sample IDs
rownames(df.p.1) <- df.p.1$SampleID

```

```{r format flie map file}
#format dataframe
m.1 <- m %>%
  mutate(across(c(SampleID, location, treatment), as.factor)) %>%
  mutate(ngDNA_100uL = dna_conc_ng_uL * 100,  # Convert to ng in 100 uL
    ngDNA_0.35g_soil = ngDNA_100uL,  # Same value, but conceptually associated with 0.35g soil
    ngDNA_per_g_soil = as.numeric(ngDNA_0.35g_soil / 0.35)) %>%  # Normalize to 1g soil
  mutate(ngDNA_per_g_soil = round(ngDNA_per_g_soil, 0))  # Round to whole numbers

#rownames as sample IDs
rownames(m.1) <- m.1$SampleID

```

## Exploratory Data Analysis  
```{r create histograms}
# Define the response variables for each dataframe
df_variables <- list(
  df.cn.1 = c("d13C", "d15N", "umol.C_g.soil", "umol.N_g.soil", "C_to_N"),
  df.p.1 = c("umol.TP_g.soil", "umol.TIP_g.soil", "umol.TOP_g.soil"),
  m.1 = c("ngDNA_per_g_soil")
)

# Create an empty list to store histogram plots
histogram_plots_list <- list()

# Loop through each dataframe and generate histograms
for (df_name in names(df_variables)) {
  df <- get(df_name)  # Retrieve dataframe by name
  
  for (var in df_variables[[df_name]]) {
    plot <- ggplot(df, aes(x = .data[[var]])) +
      geom_histogram(bins = 30, fill = "lightblue", color = "black") +
      ggtitle(paste("Histogram of", var, "in", df_name))
    
    # Store plot in the list
    histogram_plots_list[[paste(df_name, var, sep = "_")]] <- plot
    
    # Print the plot
    print(plot)
  }
}

```

### Save plots 
```{r save histograms}
# Define sub-directory paths for saving histograms (update with actual paths)
group_directories_histo <- list(
  df.cn.1 = sub.directory.2.1.2.1.fp ,
  df.p.1 = sub.directory.2.1.2.1.fp ,
  m.1 = sub.directory.2.1.2.1.fp 
)

# Iterate through each dataframe and save histogram plots
for (df_name in names(df_variables)) {
  for (var in df_variables[[df_name]]) {
    hist_key <- paste(df_name, var, sep = "_")
    
    if (!is.null(histogram_plots_list[[hist_key]])) {
      # Save histogram plots in .jpeg, .pdf, and .svg formats
      ggsave(filename = paste0(group_directories_histo[[df_name]], "/Histogram_", var, ".jpeg"),
             plot = histogram_plots_list[[hist_key]], device = "jpeg")
      ggsave(filename = paste0(group_directories_histo[[df_name]], "/Histogram_", var, ".pdf"),
             plot = histogram_plots_list[[hist_key]], device = "pdf")
      ggsave(filename = paste0(group_directories_histo[[df_name]], "/Histogram_", var, ".svg"),
             plot = histogram_plots_list[[hist_key]], device = "svg")
    }
  }
}

```

### Density plots 
```{r create density}
# Create an empty list to store density plots
density_plots_list <- list()

# Loop through each dataframe and generate density plots
for (df_name in names(df_variables)) {
  df <- get(df_name)  # Retrieve dataframe by name
  
  for (var in df_variables[[df_name]]) {
    plot <- ggplot(df, aes(x = .data[[var]])) +
      geom_density(fill = "lightblue", alpha = 0.5) +
      ggtitle(paste("Density Plot of", var, "in", df_name))
    
    # Store plot in the list
    density_plots_list[[paste(df_name, var, sep = "_")]] <- plot
    
    # Print the plot
    print(plot)
  }
}

```

### Save plots 
```{r save density}
# Define sub-directory paths for saving density plots (update with actual paths)
group_directories_density <- list(
  df.cn.1 = sub.directory.2.1.2.2.fp,
  df.p.1 = sub.directory.2.1.2.2.fp,
  m.1 = sub.directory.2.1.2.2.fp
)

# Iterate through each dataframe and save density plots
for (df_name in names(df_variables)) {
  for (var in df_variables[[df_name]]) {
    density_key <- paste(df_name, var, sep = "_")
    
    if (!is.null(density_plots_list[[density_key]])) {
      # Save density plots in .jpeg, .pdf, and .svg formats
      ggsave(filename = paste0(group_directories_density[[df_name]], "/Density_", var, ".jpeg"),
             plot = density_plots_list[[density_key]], device = "jpeg")
      ggsave(filename = paste0(group_directories_density[[df_name]], "/Density_", var, ".pdf"),
             plot = density_plots_list[[density_key]], device = "pdf")
      ggsave(filename = paste0(group_directories_density[[df_name]], "/Density_", var, ".svg"),
             plot = density_plots_list[[density_key]], device = "svg")
    }
  }
}

```

### Skewness
```{r compute skewness coefficient}
# Function to compute skewness
compute_skewness <- function(data, variable) {
  skew_value <- skewness(data[[variable]], na.rm = TRUE)
  return(data.frame(Variable = variable, Skewness = skew_value))
}

# Create an empty list to store skewness results
skew_results_list <- list()

# Loop through each dataframe and compute skewness
for (df_name in names(df_variables)) {
  df <- get(df_name)  # Retrieve dataframe by name
  
  for (var in df_variables[[df_name]]) {
    skew_results_list[[paste(df_name, var, sep = "_")]] <- compute_skewness(df, var)
  }
}

# Combine all results into a single dataframe
skew_results_df <- bind_rows(skew_results_list)

# Separate variables into normal and skewed
normal_vars <- skew_results_df %>% filter(Skewness >= -1 & Skewness <= 1)
skewed_vars <- skew_results_df %>% filter(Skewness < -1 | Skewness > 1)

```

```{r display skewness coefficient}
# Print results
print("Variables with approximately normal skewness:")
print(normal_vars)

print("Variables with high skewness:")
print(skewed_vars)

```


```{r format skewness coefficient}
# Separate variables into normal and skewed for each dataframe
skewness_results_list <- list()

for (df_name in names(df_variables)) {
  df_subset <- skew_results_df %>% filter(Variable %in% df_variables[[df_name]])
  
  normal_vars_vec <- df_subset %>% filter(Skewness >= -1 & Skewness <= 1) %>% pull(Variable)
  skewed_vars_vec <- df_subset %>% filter(Skewness < -1 | Skewness > 1) %>% pull(Variable)

  # Store results
  skewness_results_list[[df_name]] <- list(
    normal_vars = normal_vars_vec,
    skewed_vars = skewed_vars_vec
  )
  
  # Print results
  print(paste("Variables with approximately normal skewness in", df_name, ":"))
  print(normal_vars_vec)

  print(paste("Variables with high skewness in", df_name, ":"))
  print(skewed_vars_vec)
}

```

###Kurtosis   
```{r compute kurtosis}
# Function to compute kurtosis
compute_kurtosis <- function(data, variable) {
  kurt_value <- kurtosis(data[[variable]], na.rm = TRUE)
  return(data.frame(Variable = variable, Kurtosis = kurt_value))
}

# Create an empty list to store kurtosis results
kurtosis_results_list <- list()

# Loop through each dataframe and compute kurtosis
for (df_name in names(df_variables)) {
  df <- get(df_name)  # Retrieve dataframe by name
  
  for (var in df_variables[[df_name]]) {
    kurtosis_results_list[[paste(df_name, var, sep = "_")]] <- compute_kurtosis(df, var)
  }
}

# Combine results into a single dataframe and print it
kurtosis_results_df <- bind_rows(kurtosis_results_list)
print(kurtosis_results_df)  # Ensure the dataframe is created correctly

# Separate variables into categories based on kurtosis for each dataframe
kurtosis_results_list_formatted <- list()

for (df_name in names(df_variables)) {
  df_subset <- kurtosis_results_df %>% filter(Variable %in% df_variables[[df_name]])
  
  mesokurtic_vars_vec <- df_subset %>% filter(Kurtosis >= 2.5 & Kurtosis <= 3.5) %>% pull(Variable)
  leptokurtic_vars_vec <- df_subset %>% filter(Kurtosis > 3.5) %>% pull(Variable)
  platykurtic_vars_vec <- df_subset %>% filter(Kurtosis < 2.5) %>% pull(Variable)

  # Store results
  kurtosis_results_list_formatted[[df_name]] <- list(
    mesokurtic_vars = mesokurtic_vars_vec,
    leptokurtic_vars = leptokurtic_vars_vec,
    platykurtic_vars = platykurtic_vars_vec
  )
  
  # Print results
  print(paste("Variables with approximately normal kurtosis (Mesokurtic) in", df_name, ":"))
  print(mesokurtic_vars_vec)

  print(paste("Variables with high kurtosis (Leptokurtic) in", df_name, ":"))
  print(leptokurtic_vars_vec)

  print(paste("Variables with low kurtosis (Platykurtic) in", df_name, ":"))
  print(platykurtic_vars_vec)
}

```

### Save files
```{r save skewness and kurtosis}
# Define sub-directory paths for saving results (update with actual paths)
group_directories_stats <- list(
  df.cn.1 = sub.directory.2.1.2.fp,
  df.p.1 = sub.directory.2.1.2.fp,
  m.1 = sub.directory.2.1.2.fp
)

# Merge skewness and kurtosis results
skew_kurt_results_df <- full_join(skew_results_df, kurtosis_results_df, by = "Variable")

# Iterate through each dataframe and save results
for (df_name in names(df_variables)) {
  df_subset <- skew_kurt_results_df %>% filter(Variable %in% df_variables[[df_name]])
  
  write.csv(df_subset, file = paste0(group_directories_stats[[df_name]], "/Skewness_Kurtosis_", df_name, ".csv"), row.names = FALSE)
}

```

## Outliers
### Z-scores  
```{r compute z-scores}
# Function to compute Z-scores and flag outliers
detect_outliers_z <- function(data, variable) {
  mean_val <- mean(data[[variable]], na.rm = TRUE)
  sd_val <- sd(data[[variable]], na.rm = TRUE)
  
  outliers <- data %>%
    mutate(Variable = variable,  # Include variable name in output
           Original_Value = .data[[variable]],  # Include original value
           Z_score = (data[[variable]] - mean_val) / sd_val,
           Outlier = ifelse(abs(Z_score) > 3, TRUE, FALSE)) %>%
    filter(Outlier == TRUE) %>%
    select(Variable, Original_Value, Z_score, Outlier)
  
  return(outliers)
}

# Apply function only to normally skewed variables in each dataframe
outliers_z_list <- list()
for (df_name in names(df_variables)) {
  df_subset <- skew_results_df %>% filter(Variable %in% df_variables[[df_name]])
  
  normal_vars_vec <- df_subset %>% filter(Skewness >= -1 & Skewness <= 1) %>% pull(Variable)

  df <- get(df_name)
  
  outliers_z_list[[df_name]] <- bind_rows(lapply(normal_vars_vec, function(var) detect_outliers_z(df, var)))
  
  print(paste("Z-score outliers detected for", df_name, ":"))
  print(outliers_z_list[[df_name]])
}

```

### Modified Z-scores  
```{r compute modified z-scores}
# Function to compute Modified Z-scores and flag outliers
detect_outliers_modified_z <- function(data, variable) {
  median_val <- median(data[[variable]], na.rm = TRUE)
  mad_val <- mad(data[[variable]], na.rm = TRUE)
  
  outliers <- data %>%
    mutate(Variable = variable,  # Include variable name in output
           Original_Value = .data[[variable]],  # Include original value
           Modified_Z = (0.6745 * (data[[variable]] - median_val)) / mad_val,
           Outlier = ifelse(abs(Modified_Z) > 3.5, TRUE, FALSE)) %>%
    filter(Outlier == TRUE) %>%
    select(Variable, Original_Value, Modified_Z, Outlier)
  
  return(outliers)
}

# Apply function only to skewed variables in each dataframe
outliers_modified_z_list <- list()
for (df_name in names(df_variables)) {
  df_subset <- skew_results_df %>% filter(Variable %in% df_variables[[df_name]])
  
  skewed_vars_vec <- df_subset %>% filter(Skewness < -1 | Skewness > 1) %>% pull(Variable)

  df <- get(df_name)
  
  outliers_modified_z_list[[df_name]] <- bind_rows(lapply(skewed_vars_vec, function(var) detect_outliers_modified_z(df, var)))
  
  print(paste("Modified Z-score outliers detected for", df_name, ":"))
  print(outliers_modified_z_list[[df_name]])
}

```

### IQR
```{r compute IQR x1.5}
# Function to detect outliers using IQR method
detect_outliers_IQR <- function(data, variable) {
  Q1 <- quantile(data[[variable]], 0.25, na.rm = TRUE)
  Q3 <- quantile(data[[variable]], 0.75, na.rm = TRUE)
  IQR_val <- Q3 - Q1
  lower_bound <- Q1 - 1.5 * IQR_val
  upper_bound <- Q3 + 1.5 * IQR_val

  outliers <- data %>%
    mutate(Variable = variable,  # Include variable name in output
           Outlier_Value = .data[[variable]],  # Preserve value for display
           Outlier = ifelse(Outlier_Value < lower_bound | Outlier_Value > upper_bound, TRUE, FALSE)) %>%
    filter(Outlier == TRUE) %>%
    select(Variable, Outlier_Value, Outlier)
  
  return(outliers)
}

# Apply function to each dataframe separately
outlier_IQR_results_list <- list()
for (df_name in names(df_variables)) {
  df <- get(df_name)  # Retrieve dataframe by name
  
  outlier_IQR_results_list[[df_name]] <- bind_rows(lapply(df_variables[[df_name]], function(var) detect_outliers_IQR(df, var)))
  
  print(paste("IQR-based outliers detected for", df_name, ":"))
  print(outlier_IQR_results_list[[df_name]])
}

```

### Save files
```{r save outliers}
# Define groups and corresponding response variables for each dataframe
df_groups <- list(
  df.cn.1 = df_variables$df.cn.1,
  df.p.1 = df_variables$df.p.1,
  m.1 = df_variables$m.1
)

# Example sub-directory paths for each dataframe (update with actual file paths)
group_directories_outlier <- list(
  df.cn.1 = sub.directory.2.1.2.3.fp,
  df.p.1 = sub.directory.2.1.2.3.fp,
  m.1 = sub.directory.2.1.2.3.fp
)

# Iterate through each dataframe and collect outlier results
for (df_name in names(df_groups)) {
  # Extract outlier data from all three methods
  z_outliers <- outliers_z_list[[df_name]]
  modified_z_outliers <- outliers_modified_z_list[[df_name]]
  iqr_outliers <- outlier_IQR_results_list[[df_name]]
  
  # Combine the three outlier detection results
  combined_outliers <- bind_rows(
    mutate(z_outliers, Method = "Z-score"),
    mutate(modified_z_outliers, Method = "Modified Z-score"),
    mutate(iqr_outliers, Method = "IQR x1.5")
  )
  
  # Check if there are detected outliers before saving
  if (!is.null(combined_outliers) && nrow(combined_outliers) > 0) {
    file_path <- paste0(group_directories_outlier[[df_name]], "/combined_outliers_", df_name, ".csv")
    
    # Save the combined outlier data as a CSV file
    write.csv(combined_outliers, file = file_path, row.names = FALSE)
    
    print(paste("Saved combined outliers for", df_name, "to:", file_path))
  } else {
    print(paste("No outliers detected for", df_name, "- no file saved."))
  }
}

```

### Boxplots
```{r create boxplots}
# Loop through each dataframe and generate boxplots
for (df_name in names(df_variables)) {
  df <- get(df_name)  # Retrieve dataframe by name
  df$row_id <- rownames(df)  # Add row names as an identifier
  
  for (var in df_variables[[df_name]]) {
    
    # Create the plot
    p <- ggplot(df, aes(x = "", y = .data[[var]])) +
      geom_boxplot(outlier.colour = "red", outlier.shape = 16) +
      geom_text(aes(label = ifelse(.data[[var]] %in% boxplot.stats(.data[[var]])$out, row_id, "")), 
                hjust = -0.2, vjust = 0.5, size = 3, color = "red") +
      theme_bw() +
      labs(title = paste("Boxplot of", var, "in", df_name),
           x = "",
           y = "Value") +
      theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
    
    # Print the plot
    print(p)
  }
}

```

### Save plots
```{r save boxplots}
# Example sub-directory paths for each dataframe (update with actual file paths)
group_directories_boxplot <- list(
  df.cn.1 = sub.directory.2.1.2.3.fp,
  df.p.1 = sub.directory.2.1.2.3.fp,
  m.1 = sub.directory.2.1.2.3.fp
)

# Loop through each dataframe and generate boxplots
for (df_name in names(df_variables)) {
  df <- get(df_name)  # Retrieve dataframe by name
  df$row_id <- rownames(df)  # Add row names as an identifier
  
  for (var in df_variables[[df_name]]) {
    
    # Create the plot
    p <- ggplot(df, aes(x = "", y = .data[[var]])) +
      geom_boxplot(outlier.colour = "red", outlier.shape = 16) +
      geom_text(aes(label = ifelse(.data[[var]] %in% boxplot.stats(.data[[var]])$out, row_id, "")), 
                hjust = -0.2, vjust = 0.5, size = 3, color = "red") +
      theme_bw() +
      labs(title = paste("Boxplot of", var, "in", df_name),
           x = "",
           y = "Value") +
      theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

    # Print the plot
    print(p)

    # Define file paths for different formats
    file_jpeg <- paste0(group_directories_boxplot[[df_name]], "/boxplot_", var, ".jpeg")
    file_pdf  <- paste0(group_directories_boxplot[[df_name]], "/boxplot_", var, ".pdf")
    file_svg  <- paste0(group_directories_boxplot[[df_name]], "/boxplot_", var, ".svg")

    # Save the plots in different formats
    ggsave(file_jpeg, plot = p, device = "jpeg", width = 6, height = 4, units = "in", dpi = 300)
    ggsave(file_pdf, plot = p, device = "pdf", width = 6, height = 4, units = "in")
    ggsave(file_svg, plot = p, device = "svg", width = 6, height = 4, units = "in")

  }
}
```

## Patterns    
### Scatterplots  
```{r create scatterplots}
# Initialize a list to store scatterplots
scatter_plots_list <- list()

# Get all variables across the three dataframes (excluding SampleID)
all_variables <- setdiff(unlist(df_variables), "SampleID")

# Iterate through each variable as the reference (x-axis)
for (ref_var in all_variables) {
  # Identify the dataframe containing the reference variable
  df_x_name <- names(df_variables)[sapply(df_variables, function(vars) ref_var %in% vars)]
  
  if (length(df_x_name) > 0) {
    df_x <- get(df_x_name)  # Retrieve dataframe containing the reference variable
    
    # Ensure SampleID exists for merging
    if ("SampleID" %in% colnames(df_x)) {
      df_x <- df_x %>% mutate(row_id = SampleID)
    } else {
      df_x <- df_x %>% mutate(row_id = rownames(df_x))
    }
    
    # Iterate through all other variables as y-axis variables
    for (df_compare in names(df_variables)) {
      df_y <- get(df_compare)  # Retrieve dataframe containing the comparison variables
      
      for (comp_var in df_variables[[df_compare]]) {
        # Ensure we don't compare the variable against itself
        if (ref_var != comp_var) {
          # Ensure SampleID exists for merging
          if ("SampleID" %in% colnames(df_y)) {
            df_y <- df_y %>% mutate(row_id = SampleID)
          } else {
            df_y <- df_y %>% mutate(row_id = rownames(df_y))
          }
          
          # Merge the two dataframes based on row_id
          if (ref_var %in% colnames(df_x) && comp_var %in% colnames(df_y)) {
            df_merged <- df_x %>%
              select(row_id, x_value = .data[[ref_var]]) %>%
              inner_join(df_y %>% select(row_id, y_value = .data[[comp_var]]), by = "row_id") %>%
              filter(!is.na(x_value) & !is.na(y_value))

            # Calculate sample size
            n <- nrow(df_merged)

            # Generate scatterplot
            scatter_plot <- ggplot(df_merged, aes(x = x_value, y = y_value)) +
              geom_point(size = 4, shape = 1) +
              ggtitle(paste(comp_var, "vs", ref_var, "(n =", n, ")")) +
              labs(x = ref_var, y = comp_var) +
              theme_bw() +
              theme(
                plot.title = element_text(size = 18),
                axis.title.y = element_text(size = 20),
                axis.title.x = element_text(size = 20),
                axis.text.x = element_text(size = 18),
                axis.text.y = element_text(size = 18),
                strip.text.x = element_text(size = 30)
              )

            # Store plot in the list using `{ref_var}_vs_{comp_var}`
            scatter_plots_list[[paste(ref_var, "vs", comp_var, sep = "_")]] <- scatter_plot
          }
        }
      }
    }
  }
}

# Function to display scatterplots for a given reference variable
display_scatterplots <- function(ref_var) {
  selected_plots <- scatter_plots_list[grepl(paste0(ref_var, "_vs_"), names(scatter_plots_list))]
  
  # Print the selected scatterplots
  for (plot_name in names(selected_plots)) {
    print(selected_plots[[plot_name]])
  }
}

```

```{r display d13c}
#scatterplots for "d13C"
display_scatterplots("d13C")

```

```{r display d15C}
#display
display_scatterplots("d15N")

```

```{r display TOC}
#display
display_scatterplots("umol.C_g.soil")

```

```{r display TN}
#display
display_scatterplots("umol.N_g.soil")

```

```{r display C to N}
#display
display_scatterplots("C_to_N")

```

```{r display TP}
#display
display_scatterplots("umol.TP_g.soil")

```

```{r display TIP}
#display
display_scatterplots("umol.TIP_g.soil")

```

```{r display TOP}
#display
display_scatterplots("umol.TOP_g.soil")

```

```{r display DNA conc}
#display
display_scatterplots("ngDNA_per_g_soil")

```

###Save plots
```{r save scatterplots}
# Example sub-directory paths for each dataframe (update with actual file paths)
group_directories_plot <- list(
  df.cn.1 = sub.directory.2.1.2.4.fp,
  df.p.1 = sub.directory.2.1.2.4.fp,
  m.1 = sub.directory.2.1.2.4.fp
)

# Loop through each dataframe and generate plots
for (df_name in names(df_variables)) {
  df <- get(df_name)  # Retrieve dataframe by name
  df$row_id <- rownames(df)  # Add row names as an identifier
  
  for (var in df_variables[[df_name]]) {
    # --- Generate Scatterplots ---
    scatterplot_list <- scatter_plots_list[grepl(paste0(var, "_vs_"), names(scatter_plots_list))]

    # Define file paths for different formats
    directory <- group_directories_plot[[df_name]]

    # Save scatterplots for this variable
    for (scatter_name in names(scatterplot_list)) {
      scatter_p <- scatterplot_list[[scatter_name]]
      
      # Define file paths
      file_scat_jpeg <- paste0(directory, "/scatterplot_", scatter_name, ".jpeg")
      file_scat_pdf  <- paste0(directory, "/scatterplot_", scatter_name, ".pdf")
      file_scat_svg  <- paste0(directory, "/scatterplot_", scatter_name, ".svg")

      # Save scatterplots in multiple formats
      ggsave(file_scat_jpeg, plot = scatter_p, device = "jpeg", width = 6, height = 4, units = "in", dpi = 300)
      ggsave(file_scat_pdf, plot = scatter_p, device = "pdf", width = 6, height = 4, units = "in")
      ggsave(file_scat_svg, plot = scatter_p, device = "svg", width = 6, height = 4, units = "in")

    }
  }
}

```

### Save files
```{r save dataframes}
# Define the filepath for saving CSVs
save_directory <- sub.directory.2.1.1.fp

# List of dataframes to save
dataframes_to_save <- list(
  df.cn.1 = df.cn.1,
  df.p.1 = df.p.1,
  m.1 = m.1
)

# Iterate through dataframes and save as CSV
for (df_name in names(dataframes_to_save)) {
  file_path <- paste0(save_directory, "/", df_name, ".csv")
  
  # Save dataframe as CSV
  write.csv(dataframes_to_save[[df_name]], file = file_path, row.names = FALSE)
  
}

```

## End of Script  