---
title: "McMurdo Dry Valleys Nutrient Limitation"
subtitle: "  \n Growth as Surface Cover"
author: "Adam J. Solon"
date: "`r Sys.Date()`"
#output: html_document
output: 
  pdf_document:
    toc: TRUE
    fig_width: 7
    fig_height: 6
    fig_caption: true
    keep_md: true
fontsize: 12pt
#editor_options: 
#  chunk_output_type: console
---
  
# Script Summary  
This script plots surface growth in response to nutrient additions of nitrogen and phosphorus in a factorial design, with sediments collected from mountain slopes in Bonney, Hoare, and Fryxell basins in Taylor Valley of the McMurdo Dry Valleys, Antarctica. Once a week a dissecting microscope at 45x magnification was used to identify the growth of photoautotrophs on the surface of sediments that had been treated with nitrogen, phosphorus, nitrogen & phosphorus, or water-only controls. This script additionally tests how nutrients explained growth (as maximum surface cover).

### Steps of this pipeline:  
1.  Create and organize directories  
2.  Load R packages  
3.  Input files  
4.  Format Files  
5.  Summary Statistics - Max Response  
    a. Calculate average (mean) weekly growth per treatment  
    b. save files  
    c. Plots  
    d. save plots   
6.  Inferential Statistics   
    a. Transformation    
    b. GLM  
       i. Fit model - Max Response      
       ii. Residual Analysis    
       iii. Post hoc tests  
    c. GLMM  
       i. Fit model - Repeated Measures     
       ii. Residual Analysis    
       iii. Post hoc tests  

```{r echo = FALSE, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(include = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

```

```{r, echo = FALSE, include = FALSE}
# Change identifiers to your system and file naming. 
user <- "F:"
directory <- "/Projects"
project <- "/CryoHoles/Studies"
study <- "/Soils-Nutrient-Limitation-Growth_Solon-et-al"

sub.directory.2 <- "/analyses"
sub.directory.2.1 <- "/growth"
sub.directory.2.1.1 <- "/input"
sub.directory.2.1.2 <- "/surface_cover"
sub.directory.2.1.2.1 <- "/stats"
sub.directory.2.1.2.1.1 <- "/summary"
sub.directory.2.1.2.1.2 <- "/inferential"
sub.directory.2.1.2.1.2.1 <- "/models"
sub.directory.2.1.2.1.2.2 <- "/diagnostics"
sub.directory.2.1.2.2 <- "/plots"

```

### Set pathways and create directories  
```{r set paths for pipeline}
# First define the project and project directories. 
# Create pathway for pipeline
###################################################
path.fp <- paste0(user, directory, project, study)
if (!dir.exists(path.fp)) dir.create(path.fp)

```

```{r set paths for analyses}
###################################################
sub.directory.2.fp <- paste0(path.fp, sub.directory.2)
if (!dir.exists(sub.directory.2.fp)) dir.create(sub.directory.2.fp)

# Create sub-directory  growth
###################################################
sub.directory.2.1.fp <- paste0(sub.directory.2.fp, sub.directory.2.1)
if (!dir.exists(sub.directory.2.1.fp)) dir.create(sub.directory.2.1.fp)

# Create sub-directory  input
###################################################
sub.directory.2.1.1.fp <- paste0(sub.directory.2.1.fp, sub.directory.2.1.1)
if (!dir.exists(sub.directory.2.1.1.fp)) dir.create(sub.directory.2.1.1.fp)

# Create sub-directory  surface cover
###################################################
sub.directory.2.1.2.fp <- paste0(sub.directory.2.1.fp, sub.directory.2.1.2)
if (!dir.exists(sub.directory.2.1.2.fp)) dir.create(sub.directory.2.1.2.fp)

# Create sub-directory  stats
###################################################
sub.directory.2.1.2.1.fp <- paste0(sub.directory.2.1.2.fp, sub.directory.2.1.2.1)
if (!dir.exists(sub.directory.2.1.2.1.fp)) dir.create(sub.directory.2.1.2.1.fp)

# Create sub-directory  summary
###################################################
sub.directory.2.1.2.1.1.fp <- paste0(sub.directory.2.1.2.1.fp, sub.directory.2.1.2.1.1)
if (!dir.exists(sub.directory.2.1.2.1.1.fp)) dir.create(sub.directory.2.1.2.1.1.fp)

# Create sub-directory  inferential
###################################################
sub.directory.2.1.2.1.2.fp <- paste0(sub.directory.2.1.2.1.fp, sub.directory.2.1.2.1.2)
if (!dir.exists(sub.directory.2.1.2.1.2.fp)) dir.create(sub.directory.2.1.2.1.2.fp)

# Create sub-directory  models
###################################################
sub.directory.2.1.2.1.2.1.fp <- paste0(sub.directory.2.1.2.1.2.fp, sub.directory.2.1.2.1.2.1)
if (!dir.exists(sub.directory.2.1.2.1.2.1.fp)) dir.create(sub.directory.2.1.2.1.2.1.fp)

# Create sub-directory  diagnostics
###################################################
sub.directory.2.1.2.1.2.2.fp <- paste0(sub.directory.2.1.2.1.2.fp, sub.directory.2.1.2.1.2.2)
if (!dir.exists(sub.directory.2.1.2.1.2.2.fp)) dir.create(sub.directory.2.1.2.1.2.2.fp)

# Create sub-directory  plots 
###################################################
sub.directory.2.1.2.2.fp <- paste0(sub.directory.2.1.2.fp, sub.directory.2.1.2.2)
if (!dir.exists(sub.directory.2.1.2.2.fp)) dir.create(sub.directory.2.1.2.2.fp)

```

### Load R packages  
```{r Install and load packages, echo = FALSE, include = FALSE}
# install.packages("knitr")
# install.packages("kableExtra")
# install.packages("tidyverse")
# install.packages("dplyr")
# install.packages("patchwork")
# install.packages("glmmTMB")
# install.packages("DHARMa")
# install.packages("emmeans")

library(knitr); packageVersion("knitr")
library(kableExtra); packageVersion("kableExtra")
library(tidyverse); packageVersion("tidyverse")
library(dplyr); packageVersion("dplyr")
library(patchwork); packageVersion("patchwork")
library(glmmTMB); packageVersion("glmmTMB")
library(DHARMa); packageVersion("DHARMa")
library(emmeans); packageVersion("emmeans")

```

* r version: `r getRversion()`
* RStudio version: `r rstudioapi::versionInfo()$version`
* r packages:  
  knitr, `r packageVersion("knitr")` 
  kableExtra, `r packageVersion("kableExtra")` 
  tidyverse, `r packageVersion("tidyverse")` 
  dplyr, `r packageVersion("dplyr")` 
  patchwork, `r packageVersion("patchwork")` 
  glmmTMB, `r packageVersion("glmmTMB")` 
  DHARMa, `r packageVersion("DHARMa")` 
  emmeans, `r packageVersion("emmeans")` 
   

### Input Files
Required input files:   
  
1.  Spreadsheets with percent photoautotrophic surface cover for each sample from each basin. From R markdown file: 
MDV-nutrient-limitation-soils-microscope-counts.Rmd  

```{r input data files}
# Count Files
countFile.1.fp <- paste0(sub.directory.2.1.1.fp, "/BonneyBasin_counts_percent.csv") 
countFile.2.fp <- paste0(sub.directory.2.1.1.fp, "/HoareBasin_counts_percent.csv") 
countFile.3.fp <- paste0(sub.directory.2.1.1.fp, "/FryxellBasin_counts_percent.csv") 

#input data
df.b <- read.csv(countFile.1.fp, header = T)
df.h <- read.csv(countFile.2.fp, header = T)
df.f <- read.csv(countFile.3.fp, header = T)

```

# Summary Statistics  
## Calculate average growth per week per treatment group  
```{r calculate averages}
# Remove rows with sampleID BNM_11. This plate was overwatered at the beginning of the experiment and thus subjected to a different condition that none of the other samples incurred.
df.b.1 <- df.b %>% filter(SampleID != "BNM_11")

# Function to calculate summary statistics
summarize_primary_cover <- function(df) {
  df %>%
    group_by(treatment, days) %>%
    summarize(
      mean = mean(primary_cover, na.rm = TRUE),
      sd = sd(primary_cover, na.rm = TRUE),
      se = sd / sqrt(n()),
      ci_lower = mean - qt(0.975, df = n()-1) * se,
      ci_upper = mean + qt(0.975, df = n()-1) * se,
      .groups = "drop"
    ) %>%
    arrange(treatment, days)
}

# List of data frames with custom names
dfs <- list(df.b.1, df.h, df.f) %>%
  setNames(c("BonneyBasin", "HoareBasin", "FryxellBasin"))

# Apply the function to all dataframes and print as a kableExtra table
map(dfs, function(df) {
  summary_df <- summarize_primary_cover(df)
  print(kable(summary_df, format = "html") %>% kable_styling("striped", full_width = FALSE))
})

```

### Save files
```{r save sum stats}
# Ensure directories exist before saving
ensure_directory <- function(dir) {
  if (!dir.exists(dir)) {
    dir.create(dir, recursive = TRUE)  # Creates directory if it doesn’t exist
  }
}

# Function to save summary statistics as CSV
save_csv <- function(summary_df, basin_name) {
  
  # Save summary table as CSV
  write.csv(summary_df, file = file.path(sub.directory.2.1.2.1.1.fp, paste0(basin_name, "_summary.csv")), row.names = FALSE)
}

# Compute summary statistics and save CSVs
walk2(dfs, names(dfs), ~ save_csv(summarize_primary_cover(.x), .y))

```

## Plot average growth as Time Series  
```{r plot surface cover}
# Custom fill color mappings for each dataset
custom_fill <- list(
  BonneyBasin = c("CTRL" = "tan", "N" = "tan", "P" = "#01665E", "NP" = "#01665E"),
  HoareBasin = c("CTRL" = "tan", "N" = "#01665E", "P" = "tan", "NP" = "#01665E"),
  FryxellBasin = c("CTRL" = "tan", "N" = "tan", "P" = "#01665E", "NP" = "#01665E")
)

# Function to create scatterplot for each dataset
plot_percent_cover <- function(df, summary_df, title, basin_name) {
  
  # Set x-axis limits and breaks dynamically
  x_limits <- if (basin_name == "BonneyBasin") c(36, 71) else c(29, 64)
  x_breaks <- seq(x_limits[1], x_limits[2], by = 7)
  
  ggplot(df, aes(x = days, y = primary_cover)) +
    geom_jitter(aes(fill = factor(treatment, levels = c("CTRL", "N", "P", "NP"))), 
                size = 3, shape = 21, color = "black", width = 0.1, alpha = 0.35) +  # Reduced jitter point size
    geom_pointrange(data = summary_df, 
                  aes(x = days, y = mean, ymin = mean - se, ymax = mean + se, fill = factor(treatment, levels = c("CTRL", "N", "P", "NP"))), 
                  size = 1.5, shape = 21, color = "black", stroke = 0.5, alpha = 0.35) +  # Black outline added to pointrange circles
    geom_line(data = summary_df, aes(x = days, y = mean, group = treatment), 
              linetype = "dashed", color = "black", linewidth = 1) +  # **Updated `linewidth` instead of `size`**
    labs(title = title, x = "Days since Treatment", y = "Surface Cover (%)") +
    scale_y_continuous(expand = c(0.0, 0.0), limits = c(0, 100), breaks = seq(0, 100, by = 20)) +
    scale_x_continuous(limits = x_limits, breaks = x_breaks) +  # **Dynamic x-axis**
    scale_fill_manual(values = custom_fill[[basin_name]]) +  # Apply custom fill colors for jitter points
    scale_color_manual(values = custom_fill[[basin_name]]) +  # Apply custom colors for pointrange
    facet_wrap(~ factor(treatment, levels = c("CTRL", "N", "P", "NP")), nrow = 1, ncol = 4) +  # Ordered facets
    guides(color = "none", fill = "none") +
    theme_bw() +
    theme(plot.title = element_text(size = rel(1.8)),  # **Shrank title font size**
          axis.title.y = element_text(size = 18), 
          axis.title.x = element_text(size = 18),
          axis.text.x = element_text(size = 12),
          axis.text.y = element_text(size = 16),
          strip.text.x = element_text(size = 13),  # Shrink facet title font size
          axis.ticks.x.top = element_blank(),
          axis.text.x.top = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.y = element_blank())
}

# Apply the plotting function separately for each dataset
plots <- map2(dfs, names(dfs), ~ plot_percent_cover(.x, summarize_primary_cover(.x), .y, .y))

# Print each plot separately
print(plots$BonneyBasin)
print(plots$HoareBasin)
print(plots$FryxellBasin)

```

### Combine Plots for figure
```{r combine plots}
# Modify individual plots
plot_b <- plots$BonneyBasin + ggtitle("Bonney Basin") + 
  theme(axis.title.y = element_blank()) +  # Remove y-axis title for top row
  theme(plot.title = element_text(size = rel(1.3)))  # Shrink title font size

plot_h <- plots$HoareBasin + ggtitle("Hoare Basin") + 
  theme(plot.title = element_text(size = rel(1.3)))  # Shrink title font size

plot_f <- plots$FryxellBasin + ggtitle("Fryxell Basin") + 
  theme(axis.title.y = element_blank()) +  # Remove y-axis title for bottom row
  theme(plot.title = element_text(size = rel(1.3)))  # Shrink title font size

# Remove x-axis title for top and middle rows, keep for bottom
plot_b <- plot_b + theme(axis.title.x = element_blank())
plot_h <- plot_h + theme(axis.title.x = element_blank())

# Create a custom legend using ggplot
legend_plot <- ggplot() +
  geom_point(aes(x = 2.08, y = 1), shape = 21, size = 5, fill = "#01665E", color = "black", alpha = 0.35) +
  geom_point(aes(x = 2.08, y = 0), shape = 21, size = 5, fill = "tan", color = "black", alpha = 0.35) +
  annotate("text", x = 2.11, y = 1, label = "Significant growth", hjust = 0, size = 4.5) +
  annotate("text", x = 2.11, y = 0, label = "Non-significant growth", hjust = 0, size = 4.5) +
  xlim(0.8, 2.5) + ylim(-0.5, 1.5) +
  theme_void()

#combine all plots
combined_plot <- (legend_plot / plot_b / plot_h / plot_f) + 
  plot_layout(heights = c(0.5, 3, 3, 3))  # Adjust the last item's height

# Display the combined plot
print(combined_plot)

```

### Save Plots
```{r save plots}
# List of file extensions
file_types <- c("jpeg", "pdf", "svg")

# Function to save plots with specific dimensions
save_plot <- function(plot, filename, width, height) {
  for (file_type in file_types) {
    file_name <- file.path(sub.directory.2.1.2.2.fp, paste0(filename, ".", file_type))
    ggsave(file_name, plot, device = file_type, width = width, height = height)
  }
}

# Iterate through list and save each individual plot with dimensions 8 x 6
for (name in names(plots)) {
  save_plot(plots[[name]], paste0(name, "_surface_cover"), width = 8, height = 6)
}

# Save the combined plot with dimensions 8.5 x 11
save_plot(combined_plot, "MDV-nutrient-limitation-soils-combined-surface-cover", width = 8.5, height = 11)

```

# Inferential Statistics  
Using surface growth (percentage) of each plate, the mean response of each treatment will be tested for differences from the control treatment for both maximum growth response and growth across time.  

### Format  
```{r format for models}
# Define a function to process formatting and transformation
format_dataframe <- function(df, dataset_name) {
  
  # Subset to max primary_cover_2 per plate_number
  df.1 <- df %>%
    group_by(plate_number) %>%
    mutate(max_cover = max(primary_cover, na.rm = TRUE)) %>%
    filter(primary_cover == max_cover) 
  
  # Subset to keep only the latest week if more than one week shares the same max cover  
  df.2 <- df.1 %>%
    group_by(plate_number) %>%
    mutate(week = as.numeric(Week)) %>%
    mutate(max_week = max(week, na.rm = TRUE)) %>% 
    filter(week == max_week) 
  
  # Convert N and P to factors with explicit levels
  df.2$N <- factor(df.2$N, levels = c(FALSE, TRUE))
  df.2$P <- factor(df.2$P, levels = c(FALSE, TRUE))
  df.2$plate_number <- factor(df.2$plate_number)

  return(df.2)
}

# Process all datasets
df.b.2 <- format_dataframe(df.b.1, "Bonney")
df.h.2 <- format_dataframe(df.h, "Hoare")
df.f.2 <- format_dataframe(df.f, "Fryxell")

```

The beta distribution is suitable for percentages or proportions (i.e., 0 to 1), which is useful for our response variable being percent surface cover from the point-intercept method. However it can not accommodate exact values of 0 or 1 (i.e., a slight transformation is required to add a small value for any exact zeroes). Also, the default link function for beta is logit.  The logit transformation is commonly used to model the mean of a response variable that is bounded between 0 and 1, as it converts proportions (which range from 0 to 1) into values that range from negative infinity to positive infinity, and approximates a linear relationship . NOTE- the coefficients are changes in the log-odds of the outcome (and not raw probability changes).  
 
### Transform response variable (add small value to exact zeroes)  
```{r format define functions}
# Define transformation function for zero values (Cribari-Neto & Zeileis, 2010)
transform.1 <- function(x) {
  (x * (length(x) - 1) + 0.01) / (length(x))
}

# Define a function to process formatting and transformation
transform_dataframe <- function(df, dataset_name) {
  
  # Apply transformation to zero values and round appropriately
  df.1 <- df %>%
  mutate(primary_cover_2_tf = round(ifelse(primary_cover_2 == 0, 0.005, primary_cover_2), 3))

  return(df.1)
}

```

```{r format for models}
#format - max responses
df.b.3 <- transform_dataframe(df.b.2, "Bonney")
df.h.3 <- transform_dataframe(df.h.2, "Hoare")
df.f.3 <- transform_dataframe(df.f.2, "Fryxell")

#format- repeated measures
df.b.1.tf <- transform_dataframe(df.b.1, "Bonney")
df.h.1.tf <- transform_dataframe(df.h, "Hoare")
df.f.1.tf <- transform_dataframe(df.f, "Fryxell")

```

```{r models define functions}
# Define function to fit max response model, print output, and save summary
#use the glmm with 'template model builder' package and function (Brooks et al. 2017)
fit_model_glm_beta <- function(df, location, filepath) {
  
  # Fit Linear mixed model
  model <- glmmTMB(primary_cover_2_tf ~ N * P,
                    data = df, 
                    family = beta_family(),
                    control=glmmTMBControl(optimizer=optim, optArgs=list(method="BFGS"))) 
    
  # Assign model globally
  assign(paste0("glmm_beta_logit_", location), model, envir = .GlobalEnv)

  # Capture and save model summary to file while printing to console
  summary_output <- capture.output(print(summary(model)))
  writeLines(summary_output, con = filepath)
  
  # Print dataset name before printing model output
  cat("\nDataset:", location, "\n")
  cat(summary_output, sep = "\n")

  return(model)
}

# List of file extensions
file_types <- c("jpeg", "pdf", "svg")

# Function to generate and save residual plots (Base R)
plot_residuals_1 <- function(model, dataset, filepath) {

  # Simulate residuals
  simulationOutput <- simulateResiduals(fittedModel = model)

  # Print dataset name before printing residual diagnostics
  cat("\nDataset:", dataset, "\n")

  # Display the plot in the console once
  plot(simulationOutput, quantreg = TRUE)

  # Save residual plots using base R's graphics devices
  for (file_type in file_types) {
    file_name <- paste0(filepath, ".", file_type)  # Proper file name construction

    # Open graphics device
    if (file_type == "jpeg") jpeg(file_name, width = 800, height = 600)
    if (file_type == "pdf") pdf(file_name)
    if (file_type == "svg") svg(file_name)

    # Generate the residual plot again for saving
    plot(simulationOutput, quantreg = TRUE)

    # Close graphics device
    dev.off()
  }
}

# Function to generate and save residual plots (Set 2)
plot_residuals_2 <- function(model, df, dataset, filepath) {
  
  # Simulate residuals
  simulationOutput <- simulateResiduals(fittedModel = model)
  
  # Print dataset name before printing residual diagnostics
  cat("\nDataset:", dataset, "\n")
  
  # Set up two-panel layout and plot residuals for N and P (only displayed once)
  par(mfrow = c(1,2))
  plotResiduals(simulationOutput, df$N)
  plotResiduals(simulationOutput, df$P)
  
  # Save residual plots using base R's graphics devices
  for (file_type in file_types) {
    file_name <- paste0(filepath, ".", file_type)  # Proper file name construction

    # Open graphics device
    if (file_type == "jpeg") jpeg(file_name, width = 800, height = 600)
    if (file_type == "pdf") pdf(file_name)
    if (file_type == "svg") svg(file_name)
    
    # Recreate residual plots for saving
    par(mfrow = c(1,2))  # Ensure both plots are saved side by side
    plotResiduals(simulationOutput, df$N)
    plotResiduals(simulationOutput, df$P)
    
    # Close graphics device
    dev.off()
  }
}

```

# Fit Models 
A generalized linear model (GLM) approach will be used to fit the maximum growth response with factor variables with two levels each- N (TRUE, FALSE) and P (TRUE, FALSE)- to represent the addition (or not) of nitrogen or phosphorus.   

## GLM - Max Response
```{r fit glm Max}
# Run models and print & save outputs
MAX_glm_beta_logit_bn <- fit_model_glm_beta(df.b.3, "Bonney", paste0(sub.directory.2.1.2.1.2.1.fp, "/Bonney_MAX_glm_beta_logit.txt"))
MAX_glm_beta_logit_hr <- fit_model_glm_beta(df.h.3, "Hoare", paste0(sub.directory.2.1.2.1.2.1.fp, "/Hoare_MAX_glm_beta_logit.txt"))
MAX_glm_beta_logit_fx <- fit_model_glm_beta(df.f.3, "Fryxell", paste0(sub.directory.2.1.2.1.2.1.fp, "/Fryxell_MAX_glm_beta_logit.txt"))

```

### Residual Analysis  
```{r glm - Max residual analysis 1}
# Generate, print & save residual plots
plot_residuals_1(MAX_glm_beta_logit_bn, "Bonney", paste0(sub.directory.2.1.2.1.2.2.fp, "/Bonney_MAX_glm_beta_logit_residuals_1"))
plot_residuals_1(MAX_glm_beta_logit_hr, "Hoare", paste0(sub.directory.2.1.2.1.2.2.fp, "/Hoare_MAX_glm_beta_logit_residuals_1"))
plot_residuals_1(MAX_glm_beta_logit_fx, "Fryxell", paste0(sub.directory.2.1.2.1.2.2.fp, "/Fryxell_MAX_glm_beta_logit_residuals_1"))

```

Investigate Bonney Resdiuals further to determine if there is a serious issue with the residuals (or not)  
```{r glm - Max residual analysis 2}
# Generate, print & save residual plots
plot_residuals_2(MAX_glm_beta_logit_bn, df.b.3, "Bonney", paste0(sub.directory.2.1.2.1.2.2.fp, "/Bonney_MAX_glm_beta_logit_residuals_2"))

```

### Post Hoc tests  
```{r post hoc define functions}
# Function to perform post hoc tests using emmeans
perform_posthoc_tests <- function(model, location) {
  emmeans_results <- emmeans(model, pairwise ~ N * P, type = "response")
  
  # Print dataset name before printing model output
  cat("\nDataset:", location, "\n")
  #cat(summary_output, sep = "\n")

  # Print results
  print(emmeans_results)
  
  return(emmeans_results)
}

#function to extract estimated marginal means  
process_emmeans <- function(emmeans_results, location) {
  # Convert estimated marginal means to a dataframe
  emmeans_df <- as.data.frame(emmeans_results$emmeans)
  
  # Verify available column names for selection
  expected_columns <- c("N", "P", "response", "SE", "df", "asymp.LCL", "asymp.UCL")
  available_columns <- intersect(expected_columns, names(emmeans_df))
  
  # Select only available columns
  emmeans_df <- emmeans_df %>% dplyr::select(all_of(available_columns))
  
  # Round significant digits
  emmeans_df <- emmeans_df %>%
    mutate(response = round(response, 2),
           SE = round(SE, 2),
           asymp.LCL = round(asymp.LCL, 2),
           asymp.UCL = round(asymp.UCL, 2))

  # Add location as the first column
  emmeans_df <- emmeans_df %>%
    mutate(location = location) %>%
    relocate(location)  # Moves location to the first column
  
  return(emmeans_df)
}

#function to extract post hoc Tukey contrasts
process_contrasts <- function(emmeans_results, location) {
  # Convert contrasts to a dataframe
  contrasts_df <- as.data.frame(emmeans_results$contrasts)
  
  # Relabel contrast groups properly
  contrasts_df <- contrasts_df %>%
    mutate(contrast = case_when(
      contrast == "FALSE FALSE / TRUE FALSE" ~ "CTRL-N",
      contrast == "FALSE FALSE / FALSE TRUE" ~ "CTRL-P",
      contrast == "FALSE FALSE / TRUE TRUE"  ~ "CTRL-NP",
      contrast == "TRUE FALSE / FALSE TRUE"  ~ "N-P",
      contrast == "TRUE FALSE / TRUE TRUE"   ~ "N-NP",
      contrast == "FALSE TRUE / TRUE TRUE"   ~ "P-NP"
    ))

  # Verify available column names for selection
  expected_columns <- c("contrast", "odds.ratio", "SE", "p.value")
  available_columns <- intersect(expected_columns, names(contrasts_df))

  # Select only available columns
  contrasts_df <- contrasts_df %>% dplyr::select(all_of(available_columns))

  # Round significant digits
  contrasts_df <- contrasts_df %>%
    mutate(odds.ratio = round(odds.ratio, 2),  
           SE = round(SE, 2),
           p.value = formatC(signif(p.value, 3), format = "f", digits = 4),
           sig = ifelse(as.numeric(p.value) > 0.05, "ns", "*"))  # Adding sig column

  # Add location as the first column using relocate()
  contrasts_df <- contrasts_df %>%
    mutate(location = location) %>%
    relocate(location)  # Moves location to the first column

  return(contrasts_df)
}

```

```{r post hoc emmeans tukey MAX}
# Run post hoc tests for each model
posthoc_MAX_bn <- perform_posthoc_tests(MAX_glm_beta_logit_bn, "Bonney")
posthoc_MAX_hr <- perform_posthoc_tests(MAX_glm_beta_logit_hr, "Hoare")
posthoc_MAX_fx <- perform_posthoc_tests(MAX_glm_beta_logit_fx, "Fryxell")

```

### Summarize estimated marginal means  
```{r summarize emmeans means}
# Run function on each dataset
means_MAX_bn <- process_emmeans(posthoc_MAX_bn, "Bonney")
means_MAX_hr <- process_emmeans(posthoc_MAX_hr, "Hoare")
means_MAX_fx <- process_emmeans(posthoc_MAX_fx, "Fryxell")

# Combine results
final_means_MAX_df <- bind_rows(means_MAX_bn, means_MAX_hr, means_MAX_fx)

# Save dataframe as CSV
write_csv(final_means_MAX_df, paste0(sub.directory.2.1.2.1.2.1.fp, "/MDV-nutrient-limitation-soils-MAX-emmeans-means.csv"))

# Print dataframe to console
print(final_means_MAX_df)

```

### Summarize Tukey contrasts     
```{r summarize emmeans contrasts}
# Run function on each dataset
contrasts_MAX_bn <- process_contrasts(posthoc_MAX_bn, "Bonney")
contrasts_MAX_hr <- process_contrasts(posthoc_MAX_hr, "Hoare")
contrasts_MAX_fx <- process_contrasts(posthoc_MAX_fx, "Fryxell")

# Combine results
final_contrasts_MAX_df <- bind_rows(contrasts_MAX_bn, contrasts_MAX_hr, contrasts_MAX_fx)

# Save dataframe as CSV
write_csv(final_contrasts_MAX_df, paste0(sub.directory.2.1.2.1.2.1.fp, "/MDV-nutrient-limitation-soils-MAX-emmeans-contrasts.csv"))

# Print dataframe to console
print(final_contrasts_MAX_df)

```

## GLMM - Repeated Measures  
A generalized linear mixed model (GLMM) approach will be used to test for differences in growth across the time series and will be fit with fixed effects (i.e. treatments) as factor variables with two levels each- N (TRUE, FALSE) and P (TRUE, FALSE)- to represent the addition (or not) of nitrogen or phosphorus.  Additionally, a random effect of plate number will be used to account for repeated measures of individual plates across time.    
 
```{r repeated measures define functions}
# Define function to fit repeated measures model with random effects, print output, and save summary
#use the glmm with 'template model builder' package and function (Brooks et al. 2017)
fit_model_glmm_beta <- function(df, location, filepath) {
  
  # Fit Linear mixed model
  model <- glmmTMB(primary_cover_2_tf ~ N * P + (1| plate_number),
                    data = df, 
                    family = beta_family(),
                    control=glmmTMBControl(optimizer=optim, optArgs=list(method="BFGS"))) 
    
  # Assign model globally
  assign(paste0("glmm_beta_logit_", location), model, envir = .GlobalEnv)

  # Capture and save model summary to file while printing to console
  summary_output <- capture.output(print(summary(model)))
  writeLines(summary_output, con = filepath)
  
  # Print dataset name before printing model output
  cat("\nDataset:", location, "\n")
  cat(summary_output, sep = "\n")

  return(model)
}

# Function to generate and save random effects diagnostics  
plot_diagnostics_random_effects <- function(model, location, filepath) {
  # Make sure file_types is defined
  file_types <- c("jpeg", "pdf", "svg")
  
  # Print location name
  cat("\nDataset:", location, "\n")
  
  # Directly print the plot to the console
  p_console <- lme4:::dotplot.ranef.mer(ranef(model)$cond)
  print(p_console)
  
  for (file_type in file_types) {
    # Construct file name properly
    file_name <- paste0(filepath, ".", file_type)
    
    # Open the appropriate graphics device
    if (file_type == "jpeg") {
      jpeg(file_name, width = 800, height = 600)
    } else if (file_type == "pdf") {
      pdf(file_name)
    } else if (file_type == "svg") {
      svg(file_name)
    }
    
    # Generate the plot and force it to render by printing it
    p <- lme4:::dotplot.ranef.mer(ranef(model)$cond)
    print(p)
    
    # Close the graphics device
    dev.off()
  }
}

```

```{r glmm - repeated measures 1}
# repeated measures
RM_glmm_beta_logit_bn <- fit_model_glmm_beta(df.b.1.tf, "Bonney", paste0(sub.directory.2.1.2.1.2.1.fp, "/Bonney_RM_glmm_beta_logit.txt"))
RM_glmm_beta_logit_hr <- fit_model_glmm_beta(df.h.1.tf, "Hoare", paste0(sub.directory.2.1.2.1.2.1.fp, "/Hoare_RM_glmm_beta_logit.txt"))
RM_glmm_beta_logit_fx <- fit_model_glmm_beta(df.f.1.tf, "Fryxell", paste0(sub.directory.2.1.2.1.2.1.fp, "/Fryxell_RM_glmm_beta_logit.txt"))

```

### GLMM - Residual Analysis  
```{r glmm - repeated measures resid 1}
# Generate, print & save residual plots
plot_residuals_1(RM_glmm_beta_logit_bn, "Bonney", paste0(sub.directory.2.1.2.1.2.2.fp, "/Bonney_RM_glmm_beta_logit_residuals_1"))
plot_residuals_1(RM_glmm_beta_logit_hr, "Hoare", paste0(sub.directory.2.1.2.1.2.2.fp, "/Hoare_RM_glmm_beta_logit_residuals_1"))
plot_residuals_1(RM_glmm_beta_logit_fx, "Fryxell", paste0(sub.directory.2.1.2.1.2.2.fp, "/Fryxell_RM_glmm_beta_logit_residuals_1"))

```

```{r glmm - repeated measures resid 2}
# Generate, print & save residual plots
plot_residuals_2(RM_glmm_beta_logit_bn, df.b.1.tf, "Bonney", paste0(sub.directory.2.1.2.1.2.2.fp, "/Bonney_RM_glmm_beta_logit_residuals_2"))
plot_residuals_2(RM_glmm_beta_logit_hr, df.h.1.tf, "Hoare", paste0(sub.directory.2.1.2.1.2.2.fp, "/Hoare_RM_glmm_beta_logit_residuals_2"))
plot_residuals_2(RM_glmm_beta_logit_fx, df.f.1.tf, "Fryxell", paste0(sub.directory.2.1.2.1.2.2.fp, "/Fryxell_RM_glmm_beta_logit_residuals_2"))

```

```{r glmm - repeated measures random effects}
# Generate, print & save residual plots
plot_diagnostics_random_effects(RM_glmm_beta_logit_bn, "Bonney", paste0(sub.directory.2.1.2.1.2.2.fp, "/Bonney_RM_glmm_beta_logit_random_effects_diagnostics"))
plot_diagnostics_random_effects(RM_glmm_beta_logit_hr, "Hoare", paste0(sub.directory.2.1.2.1.2.2.fp, "/Hoare_RM_glmm_beta_logit_random_effects_diagnostics"))
plot_diagnostics_random_effects(RM_glmm_beta_logit_fx, "Fryxell", paste0(sub.directory.2.1.2.1.2.2.fp, "/Hoare_RM_glmm_beta_logit_random_effects_diagnostics"))

```

## Re-fit Bonney    
```{r repeated measures define functions bn}
# Define function to fit repeated measures model with random effects, print output, and save summary
#use the glmm with 'template model builder' package and function (Brooks et al. 2017)
fit_model_glmm_beta_bn <- function(df, location, filepath) {
  
  # Fit Linear mixed model
  model <- glmmTMB(primary_cover_2_tf ~ N * P + (1| plate_number),
                    dispformula = ~ N * P,
                    data = df, 
                    family = beta_family(),
                    control=glmmTMBControl(optimizer=optim, optArgs=list(method="BFGS"))) 
    
  # Assign model globally
  assign(paste0("glmm_beta_logit_", location), model, envir = .GlobalEnv)

  # Capture and save model summary to file while printing to console
  summary_output <- capture.output(print(summary(model)))
  writeLines(summary_output, con = filepath)
  
  # Print dataset name before printing model output
  cat("\nDataset:", location, "\n")
  cat(summary_output, sep = "\n")

  return(model)
}

```

```{r glmm - repeated measures bn}
# repeated measures - Bonney
RM_glmm_beta_logit_bn <- fit_model_glmm_beta_bn(df.b.1.tf, "Bonney", paste0(sub.directory.2.1.2.1.2.1.fp, "/Bonney_RM_glmm_beta_logit.txt"))

```

### Residual Analysis  
```{r glmm - repeated measures resid 1 bn}
# Generate, print & save residual plots
plot_residuals_1(RM_glmm_beta_logit_bn, "Bonney", paste0(sub.directory.2.1.2.1.2.2.fp, "/Bonney_RM_glmm_beta_logit_residuals_1"))

```

```{r glmm - repeated measures resid 2 bn}
# Generate, print & save residual plots
plot_residuals_2(RM_glmm_beta_logit_bn, df.b.1.tf, "Bonney", paste0(sub.directory.2.1.2.1.2.2.fp, "/Bonney_RM_glmm_beta_logit_residuals_2"))

```

```{r glmm - repeated measures random effects bn}
# Generate, print & save residual plots
plot_diagnostics_random_effects(RM_glmm_beta_logit_bn, "Bonney", paste0(sub.directory.2.1.2.1.2.2.fp, "/Bonney_RM_glmm_beta_logit_random_effects_diagnostics"))

```

## Re-fit Hoare  
```{r repeated measures define functions hr}
# Define function to fit repeated measures model with random effects, print output, and save summary
#use the glmm with 'template model builder' package and function (Brooks et al. 2017)  
fit_model_glmm_beta_hr <- function(df, location, filepath) {
  
  # Fit Linear mixed model
  model <- glmmTMB(primary_cover_2_tf ~ N * P + poly(days, 2) + (1 | plate_number),
                    dispformula = ~ N * P,
                    data = df, 
                    family = beta_family(),
                    control=glmmTMBControl(optimizer=optim, optArgs=list(method="BFGS")))
    
  # Assign model globally
  assign(paste0("glmm_beta_logit_", location), model, envir = .GlobalEnv)

  # Capture and save model summary to file while printing to console
  summary_output <- capture.output(print(summary(model)))
  writeLines(summary_output, con = filepath)
  
  # Print dataset name before printing model output
  cat("\nDataset:", location, "\n")
  cat(summary_output, sep = "\n")

  return(model)
}

```

```{r glmm - repeated measures hr}
# repeated measures - Bonney
RM_glmm_beta_logit_hr <- fit_model_glmm_beta_hr(df.h.1.tf, "Hoare", paste0(sub.directory.2.1.2.1.2.1.fp, "/Hoare_RM_glmm_beta_logit.txt"))

```

### Residual Analysis  
```{r glmm - repeated measures resid 1 hr}
#print & save residual plots
plot_residuals_1(RM_glmm_beta_logit_hr, "Hoare", paste0(sub.directory.2.1.2.1.2.2.fp, "/Hoare_RM_glmm_beta_logit_residuals_1"))

```

```{r glmm - repeated measures resid 2 hr}
# Generate, print & save residual plots
plot_residuals_2(RM_glmm_beta_logit_hr, df.h.1.tf, "Hoare", paste0(sub.directory.2.1.2.1.2.2.fp, "/Hoare_RM_glmm_beta_logit_residuals_2"))

```

```{r repeated measures random effects hr}
# Generate, print & save residual plots
plot_diagnostics_random_effects(RM_glmm_beta_logit_hr, "Hoare", paste0(sub.directory.2.1.2.1.2.2.fp, "/Hoare_RM_glmm_beta_logit_random_effects_diagnostics"))

```

```{r RM additional diganostics hr}
#temporal autocorrelation
simRes.hr <- simulateResiduals(RM_glmm_beta_logit_hr)

#zero inflation  
DHARMa::testZeroInflation(simRes.hr)

```

```{r additional diagnostics autocor}
simRes.hr.1 <- recalculateResiduals(simRes.hr, group = df.h.1.tf$days)

testTemporalAutocorrelation(simRes.hr.1, time = unique(df.h.1.tf$days))

```

## Re-fit Fryxell  
```{r repeated measures define functions fx}
# Define function to fit repeated measures model with random effects, print output, and save summary
#use the glmm with 'template model builder' package and function (Brooks et al. 2017)
fit_model_glmm_beta_fx <- function(df, location, filepath) {
  
  # Fit Linear mixed model
  model <- glmmTMB(primary_cover_2_tf ~ N * P + poly(days, 2) + (1| plate_number),
                   #dispformula = ~ days,
                    data = df, 
                    family = beta_family(),
                    control=glmmTMBControl(optimizer=optim, optArgs=list(method="BFGS")))
    
  # Assign model globally
  assign(paste0("glmm_beta_logit_", location), model, envir = .GlobalEnv)

  # Capture and save model summary to file while printing to console
  summary_output <- capture.output(print(summary(model)))
  writeLines(summary_output, con = filepath)
  
  # Print dataset name before printing model output
  cat("\nDataset:", location, "\n")
  cat(summary_output, sep = "\n")

  return(model)
}

```

```{r glmm - repeated measures fx}
# repeated measures - Bonney
RM_glmm_beta_logit_fx <- fit_model_glmm_beta_fx(df.f.1.tf, "Fryxell", paste0(sub.directory.2.1.2.1.2.1.fp, "/Fryxell_RM_glmm_beta_logit.txt"))

```

### Residual Analysis  
```{r glmm - repeated measures resid 1 fx}
# Generate, print & save residual plots
plot_residuals_1(RM_glmm_beta_logit_fx, "Fryxell", paste0(sub.directory.2.1.2.1.2.2.fp, "/Fryxell_RM_glmm_beta_logit_residuals_1"))

```

```{r glmm - repeated measures resid 2 fx}
# Generate, print & save residual plots
plot_residuals_2(RM_glmm_beta_logit_fx, df.f.1.tf, "Fryxell", paste0(sub.directory.2.1.2.1.2.2.fp, "/Fryxell_RM_glmm_beta_logit_residuals_2"))

```

```{r repeated measures random effects hr}
# Generate, print & save residual plots
plot_diagnostics_random_effects(RM_glmm_beta_logit_fx, "Fryxell", paste0(sub.directory.2.1.2.1.2.2.fp, "/Fryxell_RM_glmm_beta_logit_random_effects_diagnostics"))

```

### Post Hoc tests
```{r post hoc emmeans tukey RM}
# Run post hoc tests for each model
posthoc_RM_bn <- perform_posthoc_tests(RM_glmm_beta_logit_bn, "Bonney")
posthoc_RM_hr <- perform_posthoc_tests(RM_glmm_beta_logit_hr, "Hoare")
posthoc_RM_fx <- perform_posthoc_tests(RM_glmm_beta_logit_fx, "Fryxell")

```

### Summarize estimated marginal means  
```{r summarize emmeans means}
# Run function on each dataset
means_RM_bn <- process_emmeans(posthoc_RM_bn, "Bonney")
means_RM_hr <- process_emmeans(posthoc_RM_hr, "Hoare")
means_RM_fx <- process_emmeans(posthoc_RM_fx, "Fryxell")

# Combine results
final_means_RM_df <- bind_rows(means_RM_bn, means_RM_hr, means_RM_fx)

# Save dataframe as CSV
write_csv(final_means_RM_df, paste0(sub.directory.2.1.2.1.2.1.fp, "/MDV-nutrient-limitation-soils-RM-emmeans-means.csv"))

# Print dataframe to console
print(final_means_RM_df)
```

### Summarize Post Hoc tests   
```{r summarize emmeans contrasts}
# Run function on each dataset
contrasts_RM_bn <- process_contrasts(posthoc_RM_bn, "Bonney")
contrasts_RM_hr <- process_contrasts(posthoc_RM_hr, "Hoare")
contrasts_RM_fx <- process_contrasts(posthoc_RM_fx, "Fryxell")

# Combine results
final_contrasts_RM_df <- bind_rows(contrasts_RM_bn, contrasts_RM_hr, contrasts_RM_fx)

# Save dataframe as CSV
write_csv(final_contrasts_RM_df, paste0(sub.directory.2.1.2.1.2.1.fp, "/MDV-nutrient-limitation-soils-RM-emmeans-contrasts.csv"))

# Print dataframe to console
print(final_contrasts_RM_df)

```

## References  
Brooks ME, Kristensen K, van Benthem KJ, Magnusson A, Berg CW, Nielsen A, Skaug HJ, Maechler M, Bolker BM (2017). “glmmTMB Balances Speed and Flexibility Among Packages for Zero-inflated Generalized Linear Mixed Modeling.” The R Journal, 9(2), 378–400. doi:10.32614/RJ-2017-066.  

Cribari-Neto, F., & Zeileis, A. (2010). Beta regression in R. Journal of statistical software, 34, 1-24.  

# End of Script 