---
title: "McMurdo Dry Valleys Nutrient Limitation"
subtitle: "  \n Processing of Microscopic Count Data"
author: "Adam J. Solon"
date: "`r Sys.Date()`"
#output: html_document
output: 
  pdf_document:
    toc: TRUE
    fig_width: 7
    fig_height: 6
    fig_caption: true
    keep_md: true
fontsize: 12pt
#editor_options: 
#  chunk_output_type: console
---
  
# Script Summary  
This script processes a time series of surface growth of sediments collected from mountain slopes in Bonney, Hoare, and Fryxell basins in Taylor Valley  of the McMurdo Dry Valleys, Antarctica. Once a week a dissecting microscope at 45x magnification was used to identify the growth of photoautotrophs on the surface of sediments treated with nitrogen, phosphorus, nitrogen & phosphorus, and water-only controls.

### Steps of this pipeline:  
1.  Create and organize directories
2.  Load R packages
3.  Input files
4.  Format & Calculate Cover
    a. variables as factor or numeric values    
    b. remove rows with NAs. 
    c. generate boolean variables for N or P  
    d. calculate percent cover 
5.  Save files  

```{r echo = FALSE, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(include = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

```

```{r, echo = FALSE, include = FALSE}
# Change identifiers to your system and file naming. 
user <- "F:"
directory <- "/Projects"
project <- "/CryoHoles/Studies"
study <- "/Soils-Nutrient-Limitation-Growth_Solon-et-al"

sub.directory.1 <- "/data"
sub.directory.1.1 <- "/processed_files"
sub.directory.1.1.1 <- "/microscope_counts"

sub.directory.2 <- "/analyses"
sub.directory.2.1 <- "/growth"
sub.directory.2.1.1 <- "/input"

```

### Set pathways and create directories  

```{r set paths for pipeline}
# First define the project and project directories. 
# Create pathway for pipeline
###################################################
path.fp <- paste0(user, directory, project, study)
if (!dir.exists(path.fp)) dir.create(path.fp)

```

```{r set paths for processed files}
# Create sub-directory  data
###################################################
sub.directory.1.fp <- paste0(path.fp, sub.directory.1)
if (!dir.exists(sub.directory.1.fp)) dir.create(sub.directory.1.fp)

# Create sub-directory  processed files
###################################################
sub.directory.1.1.fp <- paste0(sub.directory.1.fp, sub.directory.1.1)
if (!dir.exists(sub.directory.1.1.fp)) dir.create(sub.directory.1.1.fp)

# Create sub-directory  microscope counts
###################################################
sub.directory.1.1.1.fp <- paste0(sub.directory.1.1.fp, sub.directory.1.1.1)
if (!dir.exists(sub.directory.1.1.1.fp)) dir.create(sub.directory.1.1.1.fp)

```

```{r set paths for analyses}
###################################################
sub.directory.2.fp <- paste0(path.fp, sub.directory.2)
if (!dir.exists(sub.directory.2.fp)) dir.create(sub.directory.2.fp)

# Create sub-directory  growth
###################################################
sub.directory.2.1.fp <- paste0(sub.directory.2.fp, sub.directory.2.1)
if (!dir.exists(sub.directory.2.1.fp)) dir.create(sub.directory.2.1.fp)

# Create sub-directory  input
###################################################
sub.directory.2.1.1.fp <- paste0(sub.directory.2.1.fp, sub.directory.2.1.1)
if (!dir.exists(sub.directory.2.1.1.fp)) dir.create(sub.directory.2.1.1.fp)


```

### Load R packages  
```{r Install and load packages, echo = FALSE, include = FALSE}

# install.packages("tidyverse")
# install.packages("knitr")

library(knitr); packageVersion("knitr")
library(tidyverse); packageVersion("tidyverse")
library(dplyr)

```

* r version: `r getRversion()`
* RStudio version: `r rstudioapi::versionInfo()$version`
* r packages:  
  tidyverse, `r packageVersion("tidyverse")` 
  knitr, `r packageVersion("knitr")`  

### Input Files
Required input files:   
  
1.  Spreadsheets with weekly count data for each sample from each basin.  
```{r input data files}
# Count Files
countFile.1.fp <- paste0(sub.directory.1.1.1.fp, "/MDV-soils-nutrient-limitation-microscope-counts-Bonney-basin.csv") 
countFile.2.fp <- paste0(sub.directory.1.1.1.fp, "/MDV-soils-nutrient-limitation-microscope-counts-Hoare-basin.csv") 
countFile.3.fp <- paste0(sub.directory.1.1.1.fp, "/MDV-soils-nutrient-limitation-microscope-counts-Fryxell-basin.csv") 

#input data
df.b <- read.csv(countFile.1.fp, header = T)
df.h <- read.csv(countFile.2.fp, header = T)
df.f <- read.csv(countFile.3.fp, header = T)

```

#### Format & Calculate Percent Cover  
 
1. Set column variables as factor or numeric values    
2. Remove NAs. NAs are not zeros and indicate no counts were recorded for that sample in that week (i.e., the early weeks when spot checks with only field-of-view observations were conducted before growth was detected and the point-intercept methods was used to count surface cover. And also plate 3, Bonney Basin, which was lost during experiment)  
3. Create boolean variables for N or P  
4. Calculate counts for each category as a percentage of the total surface cover  

```{r echo = FALSE, include = FALSE}
#define function
process_df <- function(df) {  
  # Convert specified columns to factor  
  df <- df %>% mutate(across(c(Year, Month, Date, basin, plate_number, SampleID, treatment, replicate, technique), as.factor))  
  
  # Convert specified columns to numeric  
  df <- df %>% mutate(across(c(Week, green_sphere, green_dot, green_branching, brown_dot, green_rope, green_patch, yellow_patch,  
                               green_filaments, fungi, moss, orange_sphere, num_counts), as.numeric))  
  
  # Add 'days' column based on 'week'  
  df <- df %>% mutate(days = 1 + (Week - 1) * 7)  
  
  # Drop missing values  
  df <- df %>% drop_na()  
  
  # Create boolean variables for each treatment  
  df <- df %>% mutate(  
    N = treatment %in% c("N", "NP"),  
    P = treatment %in% c("P", "NP")  
  )  
  
  # Reclassify boolean variables as factors  
  df <- df %>% mutate(across(c(N, P), as.factor))  
  
  # Compute percent cover variables  
  df <- df %>% mutate(  
    total = rowSums(select(., green_sphere:orange_sphere), na.rm = TRUE),  
    primary = rowSums(select(., green_sphere:moss, orange_sphere), na.rm = TRUE),  
    total_cover = (total / num_counts) * 100,  
    primary_cover = (primary / num_counts) * 100,  
    primary_cover_2 = primary / num_counts,  
    gs_cover = (green_sphere / num_counts) * 100,  
    gd_cover = (green_dot / num_counts) * 100,  
    gb_cover = (green_branching / num_counts) * 100,  
    bd_cover = (brown_dot / num_counts) * 100,  
    gr_cover = (green_rope / num_counts) * 100,  
    gp_cover = (green_patch / num_counts) * 100,  
    yp_cover = (yellow_patch / num_counts) * 100,  
    gf_cover = (green_filaments / num_counts) * 100,  
    fg_cover = (fungi / num_counts) * 100,  
    ms_cover = (moss / num_counts) * 100,  
    os_cover = (orange_sphere / num_counts) * 100  
  )  
  
  # Ensure unique weeks are sorted properly  
  unique_weeks <- sort(unique(df$Week), na.last = TRUE)  
  
  return(df)  
}  

# Apply the function to all dataframes and assign custom names  
dfs <- list(df.b, df.h, df.f) %>% setNames(c("BonneyBasin", "HoareBasin", "FryxellBasin")) %>% map(process_df)

```

```{r formtat data 2}
process_df_2 <- function(df) {  
  # Convert specified columns to factor  
  df <- df %>% mutate(across(c(Year, Month, Date, basin, plate_number, SampleID, treatment, replicate, technique), as.factor))  

  # Convert specified columns to numeric  
  df <- df %>% mutate(across(c(Week, green_sphere, green_dot, green_branching, brown_dot, green_rope, green_patch, yellow_patch,  
                               green_filaments, fungi, moss, orange_sphere, num_counts), as.numeric))  

  # Add 'days' column based on 'week'  
  df <- df %>% mutate(days = 1 + (Week - 1) * 7)  

  # Create boolean variables for each treatment  
  df <- df %>% mutate(  
    N = treatment %in% c("N", "NP"),  
    P = treatment %in% c("P", "NP")  
  )  

  # Reclassify boolean variables as factors  
  df <- df %>% mutate(across(c(N, P), as.factor))  

  # Compute percent cover variables  
  df <- df %>% mutate(  
    total = rowSums(select(., green_sphere:orange_sphere), na.rm = TRUE),  
    primary = rowSums(select(., green_sphere:moss, orange_sphere), na.rm = TRUE),  
    total_cover = (total / num_counts) * 100,  
    primary_cover = (primary / num_counts) * 100,  
    primary_cover_2 = primary / num_counts,  
    gs_cover = (green_sphere / num_counts) * 100,  
    gd_cover = (green_dot / num_counts) * 100,  
    gb_cover = (green_branching / num_counts) * 100,  
    bd_cover = (brown_dot / num_counts) * 100,  
    gr_cover = (green_rope / num_counts) * 100,  
    gp_cover = (green_patch / num_counts) * 100,  
    yp_cover = (yellow_patch / num_counts) * 100,  
    gf_cover = (green_filaments / num_counts) * 100,  
    fg_cover = (fungi / num_counts) * 100,  
    ms_cover = (moss / num_counts) * 100,  
    os_cover = (orange_sphere / num_counts) * 100  
  )  

  # Ensure unique weeks are sorted properly  
  unique_weeks <- sort(unique(df$Week), na.last = TRUE)  

  return(df)  
}

# Apply the function to all dataframes and assign custom names  
dfs.2 <- list(df.b, df.h, df.f) %>% setNames(c("BonneyBasin", "HoareBasin", "FryxellBasin")) %>% map(process_df_2)

```

```{r format 3}
#define function
replace_na_cover <- function(df) {  
  df <- df %>% mutate(across(ends_with("_cover") & !contains("primary_cover_2"), ~replace_na(.x, 0.5)))  
  return(df)  
}

#apply function
dfs.3 <- map(dfs.2, replace_na_cover)

```

### Save files
```{r save files}
# Save the processed dataframes
walk2(dfs, names(dfs), ~ saveRDS(.x, paste0(sub.directory.2.1.1.fp, "/", .y, ".rds")))
walk2(dfs, names(dfs), ~ write.csv(.x, file = paste0(sub.directory.2.1.1.fp, "/", .y, "_counts_percent.csv"), row.names = FALSE))

# Save the processed dataframes
walk2(dfs.3, names(dfs.3), ~ saveRDS(.x, paste0(sub.directory.2.1.1.fp, "/", .y, ".rds")))
walk2(dfs.3, names(dfs.3), ~ write.csv(.x, file = paste0(sub.directory.2.1.1.fp, "/", .y, "_counts_percent_time_series.csv"), row.names = FALSE))

```

##End of Script 