---
title: "McMurdo Dry Valleys Nutrient Limitation"
subtitle: "  \n Growth Response Ratios"
author: "Adam J. Solon"
date: "`r Sys.Date()`"
#output: html_document
output: 
  pdf_document:
    toc: TRUE
    fig_width: 7
    fig_height: 6
    fig_caption: true
    keep_md: true
fontsize: 12pt
#editor_options: 
#  chunk_output_type: console
---
  
# Script Summary  
This report quantifies the response of the growth of primary producers from nutrient enrichments in sediments collected from three basins in Taylor Valley of the McMurdo Dry Valleys. The ratios are comparisons of the photoautotrophic growth in treatments to the growth in controls (i.e. no nutrients added). There are four treatment groups from each basin: CTRL (controls with H2O only), N (nitrogen + H2O only), P (phosphorus + H2O only), and NP (N and P combined + H2O). Response ratios provide a quantitative effect size metric that allows for comparisons within and across basins.

### Steps of this pipeline:  
1.  Create and organize directories
2.  Load R packages
3.  Input files
4.  Format Files
5.  Compute response ratios
6.  Create plots 
7.  Save files

```{r echo = FALSE, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(include = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

```

```{r, echo = FALSE, include = FALSE}
# Change identifiers to your system and file naming. 
user <- "F:"
directory <- "/Projects"
project <- "/CryoHoles/Studies"
study <- "/Soils-Nutrient-Limitation-Growth_Solon-et-al"

sub.directory.2 <- "/analyses"
sub.directory.2.1 <- "/growth"
sub.directory.2.1.1 <- "/input"
sub.directory.2.1.2 <- "/response_ratios"
sub.directory.2.1.2.1 <- "/ratios"
sub.directory.2.1.2.2 <- "/plots"

```

### Set pathways and create directories  

```{r set paths for pipeline}
# First define the project and project directories. 
# Create pathway for pipeline
###################################################
path.fp <- paste0(user, directory, project, study)
if (!dir.exists(path.fp)) dir.create(path.fp)

```

```{r set paths for analyses}
###################################################
sub.directory.2.fp <- paste0(path.fp, sub.directory.2)
if (!dir.exists(sub.directory.2.fp)) dir.create(sub.directory.2.fp)

# Create sub-directory  growth
###################################################
sub.directory.2.1.fp <- paste0(sub.directory.2.fp, sub.directory.2.1)
if (!dir.exists(sub.directory.2.1.fp)) dir.create(sub.directory.2.1.fp)

# Create sub-directory  input
###################################################
sub.directory.2.1.1.fp <- paste0(sub.directory.2.1.fp, sub.directory.2.1.1)
if (!dir.exists(sub.directory.2.1.1.fp)) dir.create(sub.directory.2.1.1.fp)

# Create sub-directory  response ratios
###################################################
sub.directory.2.1.2.fp <- paste0(sub.directory.2.1.fp, sub.directory.2.1.2)
if (!dir.exists(sub.directory.2.1.2.fp)) dir.create(sub.directory.2.1.2.fp)

# Create sub-directory  ratios
###################################################
sub.directory.2.1.2.1.fp <- paste0(sub.directory.2.1.2.fp, sub.directory.2.1.2.1)
if (!dir.exists(sub.directory.2.1.2.1.fp)) dir.create(sub.directory.2.1.2.1.fp)

# Create sub-directory  plots 
###################################################
sub.directory.2.1.2.2.fp <- paste0(sub.directory.2.1.2.fp, sub.directory.2.1.2.2)
if (!dir.exists(sub.directory.2.1.2.2.fp)) dir.create(sub.directory.2.1.2.2.fp)

```

### Load R packages  
```{r Install and load packages, echo = FALSE, include = FALSE}
# install.packages("knitr")
# install.packages("")
# install.packages("tidyverse")
# install.packages("")
# install.packages("")
# install.packages("")
# install.packages("")

library(knitr); packageVersion("knitr")
library(kableExtra); packageVersion("kableExtra")
library(tidyverse); packageVersion("tidyverse")
library(dplyr); packageVersion("dplyr")
library(purrr); packageVersion("purrr")
library(broom); packageVersion("broom")
library(patchwork); packageVersion("patchwork")

```

* r version: `r getRversion()`
* RStudio version: `r rstudioapi::versionInfo()$version`
* r packages:  
  knitr, `r packageVersion("knitr")` 
  kableExtra, `r packageVersion("kableExtra")` 
  tidyverse, `r packageVersion("tidyverse")` 
  dplyr, `r packageVersion("dplyr")` 
  purrr, `r packageVersion("purrr")` 
  broom, `r packageVersion("broom")` 
  patchwork, `r packageVersion("patchwork")` 
   
### Input Files
Required input files:   
  
1.  Spreadsheets with percent autophototrophic surface cover for each sample from each basin.  From R markdown file: 
MDV-nutrient-limitation-soils-microscope-counts.Rmd  

```{r input}
# Count Files
countFile.1.fp <- paste0(sub.directory.2.1.1.fp, "/BonneyBasin_counts_percent.csv") 
countFile.2.fp <- paste0(sub.directory.2.1.1.fp, "/HoareBasin_counts_percent.csv") 
countFile.3.fp <- paste0(sub.directory.2.1.1.fp, "/FryxellBasin_counts_percent.csv") 

#input data
df.b <- read.csv(countFile.1.fp, header = T)
df.h <- read.csv(countFile.2.fp, header = T)
df.f <- read.csv(countFile.3.fp, header = T)

# Remove rows with sampleID BNM_11. This plate was overwatered at the beginning of the experiment and thus subjected to a different condition that none of the other samples incurred.
df.b.1 <- df.b %>% filter(SampleID != "BNM_11")

```

```{r define objects}
# Define custom treatment colors by location:
custom_treatment_colors_list <- list(
  Bonney  = c("N" = "tan",    "P" = "#01665E", "NP" = "#01665E", "CTRL" = "tan"),
  Hoare   = c("N" = "#01665E", "P" = "tan",    "NP" = "#01665E", "CTRL" = "tan"),
  Fryxell = c("N" = "#01665E",    "P" = "#01665E", "NP" = "#01665E", "CTRL" = "tan")
)

# Map shorthand dataset names to location names
locations <- c(b.1 = "Bonney", h = "Hoare", f = "Fryxell")

```

## Response Ratios by week  
### Calculations  
```{r calculations RRs - weekly}
# Initialize an empty list to store results
LRR_weekly <- list()
models_weekly <- list()

for (ds in c("b.1", "h", "f")) {
  location <- locations[ds]
  cat("\n### Processing", location, "###\n")

  df <- get(paste0("df.", ds)) %>%
    mutate(primary_cover = as.numeric(primary_cover)) %>%
    drop_na(primary_cover) %>%
    mutate(primary_cover_ln = ifelse(primary_cover > 0, log(primary_cover), 0),
           factor1 = as.factor(N),
           factor2 = as.factor(P))

  base1 <- levels(df$factor1)[1]
  base2 <- levels(df$factor2)[1]
  unique_days <- sort(unique(df$days))

  # Initialize list containers
  LRR_weekly[[location]] <- list()
  models_weekly[[location]] <- list()

  for (current_day in unique_days) {
    cat("Processing day:", current_day, "\n")

    processed_df <- df %>% filter(days == current_day)

    ctrl_ln <- processed_df %>%
      filter(factor1 == base1, factor2 == base2) %>%
      summarise(ctrl_ln = mean(primary_cover_ln, na.rm = TRUE)) %>%
      pull(ctrl_ln)

    processed_df <- processed_df %>%
      mutate(LRR = primary_cover_ln - ctrl_ln,
             treatment_label = case_when(
               factor1 == "TRUE" & factor2 == "TRUE" ~ "NP",
               factor1 == "TRUE" & factor2 != "TRUE" ~ "N",
               factor1 != "TRUE" & factor2 == "TRUE" ~ "P",
               TRUE ~ "CTRL"
             ))

    LRR_weekly[[location]][[as.character(current_day)]] <- processed_df

    wts <- processed_df %>%
      group_by(factor1, factor2) %>%
      summarise(var_ln = var(primary_cover_ln, na.rm = TRUE),
                n = n(),
                w = ifelse(is.na(var_ln) | var_ln == 0, 1, 1 / (var_ln / n)),
                .groups = "drop")

    m_wgt <- processed_df %>%
      left_join(wts, by = c("factor1", "factor2")) %>%
      lm(LRR ~ factor1 * factor2, data = ., weights = w)

    model_summary <- summary(m_wgt)
    print(model_summary)

    models_weekly[[location]][[as.character(current_day)]] <- m_wgt

    model_out_file <- file.path(sub.directory.2.1.2.1.fp,
                                paste0(location, "_", current_day, "_model_wls.txt"))
    capture.output(model_summary, file = model_out_file)
  }
}

```

### Summary  
```{r summary RRs - weekly}
# Initialize an empty list to store results
summary_results_weekly <- list()

for (location in names(LRR_weekly)) {
  summary_results_weekly[[location]] <- list()
  
  for (current_day in names(LRR_weekly[[location]])) {
    processed_df <- LRR_weekly[[location]][[current_day]]

    ctrl_ln <- processed_df %>%
      filter(treatment_label == "CTRL") %>%
      summarise(ctrl_ln = mean(primary_cover_ln, na.rm = TRUE)) %>%
      pull(ctrl_ln)

    group_summary <- processed_df %>%
      group_by(factor1, factor2, treatment_label) %>%
      summarise(mean_LRR = mean(LRR, na.rm = TRUE),
                sd_LRR = sd(LRR, na.rm = TRUE),
                n = n(),
                se_LRR = sd_LRR / sqrt(n), # standard error of the mean
                ci_low_LRR = mean_LRR - 1.96 * se_LRR, # lower 95% confidence interval
                ci_high_LRR = mean_LRR + 1.96 * se_LRR, # upper 95% confidence interval
                .groups = "drop") %>%
      mutate(fold_change = exp(mean_LRR),
             se_fold = exp(mean_LRR) * se_LRR,
             raw_mean = exp(ctrl_ln) * fold_change,
             se_raw = exp(ctrl_ln) * se_fold,
             pct_change = (fold_change - 1) * 100,
             se_pct = se_fold * 100)

    table_df <- group_summary %>%
      select(treatment_label, n, mean_LRR, se_LRR, ci_high_LRR, ci_low_LRR, raw_mean, se_raw,
             fold_change, se_fold, pct_change, se_pct) %>%
      arrange(factor(treatment_label, levels = c("CTRL", "N", "P", "NP")))

    # Store result in the list for later use
    summary_results_weekly[[location]][[current_day]] <- table_df

    # Print formatted table
    table_kable <- table_df %>%
      kable(format = "html", digits = 3,
            col.names = c("Treatment", "Sample Size",
                          "Mean LRR", "SE (LRR)", "upper 95%CI (LRR)", "lower 95%CI (LRR)",
                          "Back-transformed", "SE (Back)",
                          "Fold-Change", "SE (Fold)",
                          "Percent Change", "SE (Percent)")) %>%
      kable_styling(bootstrap_options = c("striped", "hover"))

    print(table_kable)

    # Save to CSV
    table_out_file <- file.path(sub.directory.2.1.2.1.fp,
                                paste0(location, "_weekly_", current_day, "_log_response_ratios.csv"))
    write.csv(table_df, file = table_out_file, row.names = FALSE)
  }
}
```

### Plots - Log Scale  
```{r plot RRs log - weekly}
# Initialize an empty list to store plots
forest_plots_weekly <- list()

for (location in names(summary_results_weekly)) {
  forest_plots_weekly[[location]] <- list()
  
  for (current_day in names(summary_results_weekly[[location]])) {
    # Retrieve stored summary results
    group_summary <- summary_results_weekly[[location]][[current_day]]

    # Get N and P values; assume there is exactly one row for each treatment (adjust if necessary)
    N_vals <- group_summary %>% filter(treatment_label == "N")
    P_vals <- group_summary %>% filter(treatment_label == "P")

    # Calculate interaction mean and propagate uncertainty
    interaction_mean <- N_vals$mean_LRR + P_vals$mean_LRR
    interaction_se   <- sqrt(N_vals$se_LRR^2 + P_vals$se_LRR^2)
    interaction_low  <- interaction_mean - 1.96 * interaction_se
    interaction_high <- interaction_mean + 1.96 * interaction_se

    # Generate the plot
    p1 <- ggplot(group_summary, aes(x = mean_LRR, 
                                    y = factor(treatment_label, levels = c("CTRL", "N", "P", "NP")), 
                                    fill = treatment_label)) +
      geom_point(size = 6, shape = 21, color = "black") +
      geom_errorbarh(aes(xmin = ci_low_LRR, xmax = ci_high_LRR), height = 0.2) +
      geom_vline(xintercept = 0, linewidth = 1, linetype = "dashed", color = "black") +
      geom_vline(xintercept = interaction_mean, linetype = "dotdash", color = "gray50") +
      annotate("rect",
               xmin = interaction_low,
               xmax = interaction_high,
               ymin = -Inf, ymax = Inf,
               alpha = 0.1, fill = "gray50") +
      scale_x_continuous(limits = c(-1.25, 5.1), breaks = -1:5) +
      scale_y_discrete(expand = c(0.3, 0)) +
      scale_fill_manual(values = custom_treatment_colors_list[[location]]) +
      labs(title = paste(location, "Day", current_day),
           x = "Response Ratio (log)", y = "") +
      theme_bw() +
      theme(
        plot.title = element_text(size = 25),
        axis.title.x = element_text(size = 20),
        axis.text.x = element_text(size = 22),
        axis.text.y = element_text(size = 18),
        axis.ticks.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank()
      ) +
      guides(fill = "none")

    print(p1)

    # Store plot in list for later reference
    forest_plots_weekly[[location]][[current_day]] <- p1

    # Save plots in multiple formats
    ggsave(file.path(sub.directory.2.1.2.2.fp, paste0(location, "_", current_day, "_forest_plot_LOG.jpeg")),
           plot = p1, device = "jpeg", width = 7, height = 7)
    ggsave(file.path(sub.directory.2.1.2.2.fp, paste0(location, "_", current_day, "_forest_plot_LOG.pdf")),
           plot = p1, device = "pdf", width = 7, height = 7)
    ggsave(file.path(sub.directory.2.1.2.2.fp, paste0(location, "_", current_day, "_forest_plot_LOG.svg")),
           plot = p1, device = "svg", width = 7, height = 7)
  }
}

```

###Plots - Back Transformation  
```{r plot RRs back transformed - weekly}
#initialize list
backtransform_plots_weekly <- list()

for (location in names(summary_results_weekly)) {
  backtransform_plots_weekly[[location]] <- list()
  
  for (current_day in names(summary_results_weekly[[location]])) {
    # Retrieve stored summary results
    table_df <- summary_results_weekly[[location]][[current_day]]

    # Ensure `CTRL` group's `raw_mean` is dynamically retrieved
    ctrl_raw_mean <- table_df %>%
      filter(treatment_label == "CTRL") %>%
      pull(raw_mean)

    # Compute confidence intervals while handling possible NA values
    table_df <- table_df %>%
      mutate(ci_low_raw = ifelse(is.na(se_raw), NA, raw_mean - qnorm(0.975) * se_raw),
             ci_high_raw = ifelse(is.na(se_raw), NA, raw_mean + qnorm(0.975) * se_raw))

    # Generate the plot
    p2 <- ggplot(table_df, aes(x = raw_mean, y = factor(treatment_label, 
                        levels = c("CTRL", "N", "P", "NP")), color = treatment_label)) +
      geom_point(size = 4) +
      geom_errorbarh(aes(xmin = ci_low_raw, xmax = ci_high_raw), height = 0.2, na.rm = TRUE) +  # Avoid plotting NA values
      geom_vline(xintercept = ctrl_raw_mean, linetype = "dashed", color = "gray70") +  # Dynamically set CTRL reference line
      scale_x_continuous(limits = c(-0.1, 105), breaks = seq(0, 100, 20)) +
      scale_color_manual(values = custom_treatment_colors_list[[location]]) +
      labs(title = paste(location, "Day", current_day),
           x = "Surface Cover (%)", y = "") +
      theme_bw() +
      theme(
        plot.title = element_text(size = 25),
        axis.title.x = element_text(size = 20),
        axis.text.x = element_text(size = 20),
        axis.text.y = element_text(size = 18),
        axis.ticks.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank()
      ) +
      guides(color = "none")

    print(p2)

    # Store plot in list for later reference
    backtransform_plots_weekly[[location]][[current_day]] <- p2

    # Save plots in multiple formats
    ggsave(file.path(sub.directory.2.1.2.2.fp, paste0(location, "_", current_day, "_forest_plot_back_transformed.jpeg")),
           plot = p2, device = "jpeg", width = 7, height = 5)
    ggsave(file.path(sub.directory.2.1.2.2.fp, paste0(location, "_", current_day, "_forest_plot_back_transformed.pdf")),
           plot = p2, device = "pdf", width = 7, height = 5)
    ggsave(file.path(sub.directory.2.1.2.2.fp, paste0(location, "_", current_day, "_forest_plot_back_transformed.svg")),
           plot = p2, device = "svg", width = 7, height = 5)
  }
}
```

## Maximum Response per Treatment
### Calculations  
```{r calculations RRs - max}
# Initialize list containers before processing
LRR_peak <- list()
models_peak <- list()

for (ds in c("b.1", "h", "f")) {
  location <- locations[ds]
  cat("\n### Processing", location, "###\n")

  df <- get(paste0("df.", ds)) %>%
    mutate(primary_cover = as.numeric(primary_cover)) %>%
    drop_na(primary_cover) %>%
    mutate(primary_cover_ln = ifelse(primary_cover > 0, log(primary_cover), 0),
           factor1 = as.factor(N),
           factor2 = as.factor(P))

  base1 <- levels(df$factor1)[1]
  base2 <- levels(df$factor2)[1]

  # Initialize location-specific sublists
  LRR_peak[[location]] <- list()
  models_peak[[location]] <- list()

  # Identify the peak LRR week per treatment in this location
  peak_days <- summary_results_weekly[[location]] %>%
    bind_rows(.id = "current_day") %>%
    filter(treatment_label %in% c("N", "P", "NP")) %>%
    group_by(treatment_label) %>%
    slice_max(mean_LRR, with_ties = FALSE) %>%
    select(treatment_label, current_day) %>%
    ungroup()

  # Get a single peak week to extract all primary_cover_ln values from treatments
  peak_week <- unique(peak_days$current_day)

  # Extract all relevant data from that peak week
  processed_df <- df %>% filter(days %in% as.numeric(peak_week))

  ctrl_ln <- processed_df %>%
    filter(factor1 == base1, factor2 == base2) %>%
    summarise(ctrl_ln = mean(primary_cover_ln, na.rm = TRUE)) %>%
    pull(ctrl_ln)

  processed_df <- processed_df %>%
    mutate(LRR = primary_cover_ln - ctrl_ln,
           treatment_label = case_when(
             factor1 == "TRUE" & factor2 == "TRUE" ~ "NP",
             factor1 == "TRUE" & factor2 != "TRUE" ~ "N",
             factor1 != "TRUE" & factor2 == "TRUE" ~ "P",
             TRUE ~ "CTRL"
           ))

  # Store peak week data
  LRR_peak[[location]] <- processed_df

  # Compute weights for regression
  wts <- processed_df %>%
    group_by(factor1, factor2) %>%
    summarise(var_ln = var(primary_cover_ln, na.rm = TRUE),
              n = n(),
              w = ifelse(is.na(var_ln) | var_ln == 0, 1, 1 / (var_ln / n)),
              .groups = "drop")

  # Perform weighted regression on **all treatments from peak week**
  m_wgt <- processed_df %>%
    left_join(wts, by = c("factor1", "factor2")) %>%
    lm(LRR ~ factor1 * factor2, data = ., weights = w)

  model_summary <- summary(m_wgt)
  print(model_summary)

  # Store final model (one per location)
  models_peak[[location]] <- m_wgt

  model_out_file <- file.path(sub.directory.2.1.2.1.fp,
                              paste0(location, "_peak_model_wls.txt"))
  capture.output(model_summary, file = model_out_file)
}

```

### Summary
```{r summary RRs - max}
# Initialize an empty list to store peak data
peak_values_list <- list()

for (location in names(summary_results_weekly)) {
  # Combine all weeks' data into one data frame for this location
  combined_data <- summary_results_weekly[[location]] %>%
    bind_rows(.id = "current_day")

  # Identify the week with the highest LRR for N, P, and NP across all weeks
  max_weeks <- combined_data %>%
    filter(treatment_label %in% c("N", "P", "NP")) %>%
    group_by(treatment_label) %>%
    slice_max(mean_LRR, with_ties = FALSE) %>%
    ungroup()

  # Select only the max LRR values for each treatment
  selected_data <- max_weeks %>%
    select(treatment_label, n, mean_LRR, se_LRR, ci_low_LRR, ci_high_LRR)

  # Add CTRL as a reference point at 0
  ctrl_row <- tibble(treatment_label = "CTRL", mean_LRR = 0, se_LRR = 0, ci_low_LRR = 0, ci_high_LRR = 0)
  selected_data <- bind_rows(ctrl_row, selected_data)

  # Get N and P values for interaction calculations
  N_vals <- selected_data %>% filter(treatment_label == "N")
  P_vals <- selected_data %>% filter(treatment_label == "P")

  # Calculate interaction mean and propagate uncertainty
  interaction_mean <- N_vals$mean_LRR + P_vals$mean_LRR
  interaction_se   <- sqrt(((N_vals$se_LRR^2) / N_vals$n) + ((P_vals$se_LRR^2) / P_vals$n)) #standard error of difference between means
  interaction_low  <- interaction_mean - 1.96 * interaction_se # lower 95% confidence interval
  interaction_high <- interaction_mean + 1.96 * interaction_se # upper 95% confidence interval

  # Add interaction row and rename it as "Additive"
  interaction_row <- tibble(
    treatment_label = "Additive",
    mean_LRR = interaction_mean,
    se_LRR = interaction_se,
    ci_low_LRR = interaction_low,
    ci_high_LRR = interaction_high
  )
  selected_data <- bind_rows(selected_data, interaction_row)

  # Arrange rows in the desired order
  selected_data <- selected_data %>%
    mutate(treatment_label = factor(treatment_label, levels = c("CTRL", "N", "P", "NP", "Additive"))) %>%
    arrange(treatment_label)

  # Store processed data for later reference
  peak_values_list[[location]] <- selected_data
}

# Combine all locations into one data frame
peak_values_combined <- bind_rows(peak_values_list, .id = "location")

# Print table using kableExtra
peak_values_combined %>%
  kable(format = "html") %>%
  kable_styling(full_width = FALSE)

# Save as CSV
write.csv(peak_values_combined, file.path(sub.directory.2.1.2.fp, "MDV-nutrient-limitation-soils_LRR-forest-plot-MAX-ratios.csv"), row.names = FALSE)

```

### Plots - Log Scale  
```{r plot RRs log - max}
# Initialize an empty list to store plots
forest_plots_max_weekly <- list()

for (location in names(summary_results_weekly)) {
  # Combine all weeks' data into one data frame for this location
  combined_data <- summary_results_weekly[[location]] %>%
    bind_rows(.id = "current_day")

  # Identify the week with the highest LRR for N, P, and NP across all weeks
  max_weeks <- combined_data %>%
    filter(treatment_label %in% c("N", "P", "NP")) %>%
    group_by(treatment_label) %>%
    slice_max(mean_LRR, with_ties = FALSE) %>%
    ungroup()

  # Select only the max LRR values for each treatment
  selected_data <- max_weeks %>%
    select(treatment_label, n, mean_LRR, se_LRR, ci_low_LRR, ci_high_LRR)

  # Add CTRL as a reference point at 0
  ctrl_row <- tibble(treatment_label = "CTRL", mean_LRR = 0, se_LRR = 0, ci_low_LRR = 0, ci_high_LRR = 0)
  selected_data <- bind_rows(ctrl_row, selected_data)

  # Get N and P values for interaction calculations
  N_vals <- selected_data %>% filter(treatment_label == "N")
  P_vals <- selected_data %>% filter(treatment_label == "P")

  # Calculate interaction mean and propagate uncertainty
  interaction_mean <- N_vals$mean_LRR + P_vals$mean_LRR
  interaction_se   <- sqrt(((N_vals$se_LRR^2)/N_vals$n) + ((P_vals$se_LRR^2)/P_vals$n))
  interaction_low  <- interaction_mean - 1.96 * interaction_se
  interaction_high <- interaction_mean + 1.96 * interaction_se

  # Generate the plot
  p1 <- ggplot(selected_data, aes(x = mean_LRR, 
                                  y = factor(treatment_label, levels = c("CTRL", "N", "P", "NP")), 
                                  fill = treatment_label)) +
    geom_point(size = 4, shape = 21, color = "black", alpha = 0.55) +
    geom_errorbarh(aes(xmin = ci_low_LRR, xmax = ci_high_LRR), height = 0.2) +
    geom_vline(xintercept = 0, linewidth = 1, linetype = "dashed", color = "black") +  # Fixed CTRL reference at 0
    geom_vline(xintercept = interaction_mean, linetype = "dotdash", color = "gray50") +
    annotate("rect",
             xmin = interaction_low,
             xmax = interaction_high,
             ymin = -Inf, ymax = Inf,
             alpha = 0.1, fill = "gray50") +
    scale_x_continuous(limits = c(-1.25, 5.1), breaks = -1:5) +
    scale_y_discrete(expand = c(0.3, 0)) +
    scale_fill_manual(values = custom_treatment_colors_list[[location]]) +
    labs(title = paste(location, "Peak LRR Values"),
         x = "Response Ratio (log)", y = "") +
    theme_bw() +
    theme(
      plot.title = element_text(size = 25),
      axis.title.x = element_text(size = 20),
      axis.text.x = element_text(size = 22),
      axis.text.y = element_text(size = 18),
      axis.ticks.y = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.y = element_blank()
    ) +
    guides(fill = "none")

  print(p1)

  # Store plot in list for later reference
  forest_plots_max_weekly[[location]] <- p1

  # Save plots in multiple formats
  ggsave(file.path(sub.directory.2.1.2.2.fp, paste0(location, "_forest_plot_peak_LRR.jpeg")),
         plot = p1, device = "jpeg", width = 7, height = 7)
  ggsave(file.path(sub.directory.2.1.2.2.fp, paste0(location, "_forest_plot_peak_LRR.pdf")),
         plot = p1, device = "pdf", width = 7, height = 7)
  ggsave(file.path(sub.directory.2.1.2.2.fp, paste0(location, "_forest_plot_peak_LRR.svg")),
         plot = p1, device = "svg", width = 7, height = 7)
}
```

### Plots - Back-Transformation  
```{r RRs plots back transformed - max}
# Initialize an empty list to store plots
backtrans_plots_max_weekly <- list()

for (location in names(summary_results_weekly)) {
  # Combine all weeks' data into one data frame for this location
  combined_data <- summary_results_weekly[[location]] %>%
    bind_rows(.id = "current_day")

  # Identify the week with the highest raw_mean for N, P, and NP across all weeks
  max_weeks <- combined_data %>%
    filter(treatment_label %in% c("N", "P", "NP")) %>%
    group_by(treatment_label) %>%
    slice_max(raw_mean, with_ties = FALSE) %>%
    ungroup()

  # Select only the max raw_mean values for each treatment
  selected_data <- max_weeks %>%
    select(treatment_label, raw_mean, se_raw) %>%
    mutate(ci_low_raw = raw_mean - qnorm(0.975) * se_raw,
           ci_high_raw = raw_mean + qnorm(0.975) * se_raw)

  # Ensure `CTRL` group's `raw_mean` is dynamically retrieved
  ctrl_raw_mean <- combined_data %>%
    filter(treatment_label == "CTRL") %>%
    summarise(ctrl_raw_mean = mean(raw_mean, na.rm = TRUE)) %>%
    pull(ctrl_raw_mean)

  # Add CTRL as a reference point at its computed mean
  ctrl_row <- tibble(treatment_label = "CTRL", raw_mean = ctrl_raw_mean, se_raw = 0, 
                     ci_low_raw = ctrl_raw_mean, ci_high_raw = ctrl_raw_mean)
  selected_data <- bind_rows(ctrl_row, selected_data)

  # Generate the plot
  p2 <- ggplot(selected_data, aes(x = raw_mean, y = factor(treatment_label, 
                      levels = c("CTRL", "N", "P", "NP")), color = treatment_label)) +
    geom_point(size = 4) +
    geom_errorbarh(aes(xmin = ci_low_raw, xmax = ci_high_raw), height = 0.2, na.rm = TRUE) +  # Avoid plotting NA values
    geom_vline(xintercept = ctrl_raw_mean, linetype = "dashed", color = "gray70") +  # Dynamically set CTRL reference line
    scale_x_continuous(limits = c(-0.1, 105), breaks = seq(0, 100, 20)) +
    scale_color_manual(values = custom_treatment_colors_list[[location]]) +
    labs(title = paste(location, "Peak Back-Transformed Surface Cover"),
         x = "Surface Cover (%)", y = "") +
    theme_bw() +
    theme(
      plot.title = element_text(size = 25),
      axis.title.x = element_text(size = 20),
      axis.text.x = element_text(size = 20),
      axis.text.y = element_text(size = 18),
      axis.ticks.y = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.y = element_blank()
    ) +
    guides(color = "none")

  print(p2)

  # Store plot in list for later reference
  backtrans_plots_max_weekly[[location]] <- p2

  # Save plots in multiple formats
  ggsave(file.path(sub.directory.2.1.2.2.fp, paste0(location, "_forest_plot_peak_back_transformed.jpeg")),
         plot = p2, device = "jpeg", width = 7, height = 7)
  ggsave(file.path(sub.directory.2.1.2.2.fp, paste0(location, "_forest_plot_peak_back_transformed.pdf")),
         plot = p2, device = "pdf", width = 7, height = 7)
  ggsave(file.path(sub.directory.2.1.2.2.fp, paste0(location, "_forest_plot_peak_back_transformed.svg")),
         plot = p2, device = "svg", width = 7, height = 7)
}

```

```{r combine plots}
# Load necessary library
library(patchwork)

# Modify the plots
forest_plots_modified <- forest_plots_modified <- list(
  "Bonney"  = forest_plots_max_weekly[["Bonney"]] + labs(title = "Bonney Basin") + 
              theme(plot.title = element_text(size = 18), axis.title.x = element_blank(), axis.text.x = element_blank(), axis.text.y = element_text(size = 14), ),
  "Hoare"   = forest_plots_max_weekly[["Hoare"]] + labs(title = "Hoare Basin") + 
              theme(plot.title = element_text(size = 18), axis.title.x = element_blank(), axis.text.x = element_blank(), axis.text.y = element_text(size = 14), ),
  "Fryxell" = forest_plots_max_weekly[["Fryxell"]] + labs(title = "Fryxell Basin") + 
              theme(plot.title = element_text(size = 18), axis.text.y = element_text(size = 14), axis.text.x = element_text(size = 16), axis.title.x = element_text(size = 18))
)

# Create a custom legend using ggplot
legend_plot <- ggplot() +
  geom_point(aes(x = 2.08, y = 1), shape = 21, size = 5, fill = "#01665E", color = "black", alpha = 0.55) +
  geom_point(aes(x = 2.08, y = -0.05), shape = 21, size = 5, fill = "tan", color = "black", alpha = 0.55) +
  annotate("text", x = 2.11, y = 1, label = "Significant effect", hjust = 0, size = 4.5) +
  annotate("text", x = 2.11, y = -0.05, label = "Non-significant effect", hjust = 0, size = 4.5) +
  xlim(0.8, 2.5) + ylim(-0.5, 1.5) +
  theme_void()

#combine all plots
combined_plot <- (legend_plot / forest_plots_modified[["Bonney"]] / forest_plots_modified[["Hoare"]] / forest_plots_modified[["Fryxell"]]) + 
  plot_layout(heights = c(0.7, 3, 3, 3))  # Adjust the last item's height

# Print the combined plot
print(combined_plot)

# Save combined plot in multiple formats
ggsave(file.path(sub.directory.2.1.2.fp, "MDV-nutrient-limitation-soils_LRR-forest-plot-MAX-combine.jpeg"),
       plot = combined_plot, device = "jpeg", width = 8.5, height = 8.5)

ggsave(file.path(sub.directory.2.1.2.fp, "MDV-nutrient-limitation-soils_LRR-forest-plot-MAX-combine.pdf"),
       plot = combined_plot, device = "pdf", width = 7, height = 5)

ggsave(file.path(sub.directory.2.1.2.fp, "MDV-nutrient-limitation-soils_LRR-forest-plot-MAX-combine.svg"),
       plot = combined_plot, device = "svg", width = 7, height = 7)

```

## End of Script  